### 2013-09-29 tex/derivations.tex

\documentclass{article}
\usepackage{minx_math}
\usepackage{framed, mathtools, epstopdf,subfigure}

\begin{document}

\section{No False Negatives}

Let us intuitively describe the logic.

Let $n$ be some large number and let $\hat{f}_S$ be the optimal solution with a false negative.

Let $f^*_S$ be the population optimal solution.

\begin{align*}
\hat{R}( \hat{f}_S ) &\geq R(\hat{f}_S) - gap_n \\
	&\geq R(f^*_S miss) - gap_n \\
	&\geq R(f^*_S) + gap - gap_n \\
	&\geq \hat{R}(f^*_S) + gap - 2 gap_n \\
	&\geq \hat{R}(f^*_S) + pen(f^*_S) - pen(f^*_S) + pen(\hat{f}_S) - pen(\hat{f}_S) + gap - 2 gap_n
\end{align*}

We have then that
\[
\hat{R}(\hat{f}_S) + pen(\hat{f}_S) \geq \hat{R}(f^*_S) + pen(f^*_S) + (pen(\hat{f}_S) - pen(f^*_S) + gap - 2 gap_n)
\]

Because $pen$ and $gap_n$ decreases with $n$ and $gap$ does not, for large enough $n$, we would reach a contradiction.

The only goal then is to figure out $gap_n$.

A more precisely statement is, with probability at least $1-\delta$, 
\[
\sup_{f_1, ..., f_s} | \hat{R}(f_S) - R(f_S) | \leq gap_n
\]

\subsection{Covering Number}

\section{Functions Which Are Not Additively Faithful}

\textbf{Example 1:}
\[
f(x_1, x_2) = \sin( 2\pi x_1) \sin( 2 \pi x_2) \qquad \trm{ for } (x_1, x_2) \in [0,1]^2
\]
Note that for all $x_1$, $\int_{x_2} f(x_1, x_2) d x_2 = 0$ and also, for all $x_2$, $\int_{x_1} f(x_1, x_2) d x_1 = 0$.
An additive model would set $f_1 = 0$ and $f_2 = 0$.

\begin{figure}[htp]
\subfigure[Example 1]{\includegraphics[scale=0.5]{../code/figs/sine_wave_funct.eps}}
\subfigure[Example 2]{\includegraphics[scale=0.4]{../code/figs/rising_slope_funct.eps}}
\end{figure}

\textbf{Example 2:}
\[
f(x_1, x_2) = x_1 x_2 \qquad \trm{ for } x_1 \in [-1,1], x_2 \in [0,1]
\]
Note that for all $x_2$, $\int_{x_1} f(x_1, x_2) d x_1 = 0$, therefore, we expect $f_2 = 0$ under the additive model.

This function, for every fixed $x_2$, is a zero-intercept linear function of $x_1$ with slope exactly $x_2$.

\section{Convex Functions are Additively Faithful}

Let $\mu$ be a probability measure on $C=[0,1]^s$, $f(x)$ be a multivariate function on $C$. We say that $f$ depends on coordinate $i$ if there exist $x'_i \neq x_i$ such that $f(x'_i, x_{-i})$ and $f(x_i, x_{-i})$ are different functions of $x_{-i}$. (on some measurable set)

\begin{theorem}
Let $\mu$ be a probability measure on $C=[0,1]^s$. Let $f: C \rightarrow \R$ be a convex function, twice differentiable.

Suppose $f$ depends on all coordinates. Let $f_1,...,f_s \coloneqq \arg\min \{ \E | f(X) - \sum_k f_k(X_k) |^2 \,:\, \forall k, f_k \trm{ convex }, \E f_k(X_k) = 0\}$

 Then $f_1,...,f_n$ are non-constant functions.
 \end{theorem}

 \begin{lemma}
\label{lem:int_reduction}
Let $\mu$ be a probability measure on $C=[0,1]^s$. Let $f: C \rightarrow \R$ be a convex function, twice differentiable. Suppose that $\E f(X) = 0$.

Let $\ds f^*_1,...,f^*_s \coloneqq \arg\min \{ \E | f(X) - \sum_{k=1}^s f_k(X_k) |^2 \,:\, \forall k, f_k \trm{ convex }, \E f_k(X_k) = 0\}$ \\

Then $f^*_k(x_k) = \E[ f(X) | x_k]$.
\end{lemma}

 \begin{proof}

Let $f^*_1,...,f^*_s$ be the minimizers as defined. It must be then that $f^*_k$ minimizes $\{ \E | f(X) - \sum_{k' \neq k} f^*_{k'} (X_{k'}) - f_k (X_k)|^2 \,:\, f_k \trm{ convex }, \E f_k(X_k) = 0 \}$.

Fix $x_k$, we will show that the value $\E[ f(X) | x_k]$ minimizes 
\[
\min_{ f_k(x_k) } \int_{\mathbf{x}_{-k}} p(x) | f(\mathbf{x}) - \sum_{k' \neq k} f^*_{k'} (x_{k'}) - f_k (x_k)|^2 d \mathbf{x}_{-k}.
\]

Take the derivative with respect to $f_k(x_k)$ and set it equal to zero, we get that
\begin{align*}
\int_{\mathbf{x}_{-k}} p(\mathbf{x}) f_k(x_k) d \mathbf{x}_{-k} &= \int_{\mathbf{x}_{-k}} p(\mathbf{x}) ( f(\mathbf{x}) - \sum_{k' \neq k} f^*_{k'} (x_{k'}) ) d \mathbf{x}_{-k}  \\
p(x_k) f_k(x_k) &= \int_{\mathbf{x}_{-k}} p(\mathbf{x}) f(\mathbf{x}) d \mathbf{x}_{-k} 
 \end{align*}

Now, we verify that as a function of $x_k$, $\E[ f(X) | x_k]$ has mean zero and is convex. The former is true because $\E f(X) = 0$; the latter is true because for every $\mathbf{x}_{-k}$, $f(x_k, \mathbf{x}_{-k})$ is a convex function with respect to $x_k$ and therefore, $\int_{x_{-k}} p(\mathbf{x}|x_k) f(x_k, \mathbf{x}_{-k}) d \mathbf{x}_{-k}$ is still convex.

 \end{proof}

 \begin{proposition}
Let $p$ be a probability distribution on $C=[0,1]^s$. Let $f: C \rightarrow \R$ be a convex function, twice differentiable. Suppose $p$ is a product distribution.

Let $f^*_1,...,f^*_s \coloneqq \arg\min \{ \E | f(X) - \sum_k f_k(X_k) |^2 \,:\, \forall k, f_k \trm{ convex }, \E f_k(X_k) = 0\}$.

The following are equivalent:
\begin{packed_enum}
\item $f$ does not depends on coordinate $k$
\item For all $x_k$, $\E[ f(X) | x_k] = 0$.
\end{packed_enum}
\end{proposition}

\begin{proof}

The first condition trivially implies the second because $\E f(X) = 0$.\\

Fix $k$. Suppose that, for all $x_k$, $\E[ f(X) | x_k] = 0$. \\

By the assumption that $p$ is a product measure, we know that, for all $x_k$,
\begin{align*}
p(x_k) \E[ f(X) | x_k ] &= \int_{\mathbf{x}_{-k}} p(\mathbf{x_{-k}} \,|\, x_k) f(x_k,\mathbf{x}_{-k}) d \mathbf{x}_{-k} \\
		&=  \int_{\mathbf{x}_{-k}} p(\mathbf{x_{-k}}) f(x_k,\mathbf{x}_{-k}) d \mathbf{x}_{-k} =0
\end{align*}

For every $\mathbf{x}_{-k}$, we define the derivative
\[
g(\mathbf{x}_{-k}) \coloneqq  \lim_{x_k \rightarrow 0^+} \frac{f(x_k, \mathbf{x}_{-k}) - f(0, \mathbf{x}_{-k})}{x_k}
\]
$g(\mathbf{x}_{-k})$ is well-defined by the assumption that $f$ is everywhere differentiable.\\

We now describe two facts about $g$.

Fact 1. By exchanging limit with the integral, we reason that 
\[
\int_{\mathbf{x}_{-k}} p(\mathbf{x}_{-k}) g(\mathbf{x}_{-k}) d \mathbf{x}_{-k} = 0
\]

Fact 2. Because $f$ is convex, $g(\mathbf{x}_{-k})$ is a component of the subgradient $\partial_{\mathbf{x}} f(0, \mathbf{x}_{-k})$. (the subgradient coincides with the gradient by assumption that $f$ is twice differentiable)

Therefore, using the first order characterization of a convex function, we have
\begin{align*}
f(\mathbf{x}') &\geq f( \mathbf{x}) + \partial_{\mathbf{x}} f(\mathbf{x})^\tran (\mathbf{x}' - \mathbf{x}) \quad \trm{ for all } \mathbf{x}', \mathbf{x} \\
f(x_k, \mathbf{x}_{-k}) &\geq f(0, \mathbf{x}_{-k}) + g(\mathbf{x}_{-k}) x_k \quad \trm{ for all } x_k, \mathbf{x}_{-k}
\end{align*}

Because, for all $x_k, \mathbf{x}_{-k}$, 
\[
f(x_k, \mathbf{x}_{-k}) - f(0, \mathbf{x}_{-k}) - g(\mathbf{x}_{-k}) x_k \geq 0
\]
and
\[
\int_{\mathbf{x}_{-k}} p(\mathbf{x}_{-k}) (f(x_k, \mathbf{x}_{-k}) - f(0, \mathbf{x}_{-k}) - g(\mathbf{x}_{-k}) x_k) d\mathbf{x}_{-k} = 0
\]
we conclude that for all $x_k, \mathbf{x}_{-k}$, $f(x_k, \mathbf{x}_{-k}) = f(0, \mathbf{x}_{-k}) + g(\mathbf{x}_{-k}) x_k$.

The Hessian of $f$ then (guaranteed to exist by assumption) has a zero on the $k$-th main diagonal entry.

By proposition from Horn and Johnson [TODO ref], such a matrix is positive semidefinite if and only if the $k$-th row and column are also zero. \\

Since $k$-th row and column correspond precisely to the gradient of $g(\mathbf{x}_{-k})$, we conclude that $g$ must be a constant function. It follows therefore that $g = 0$ because it integrates to 0. \\

So we have that for all $x_k, \mathbf{x}_{-k}$, $f(x_k, \mathbf{x}_{-k}) = f(0, \mathbf{x}_{-k})$, which concludes our proof.

\end{proof}

\section{Technical Details}

\subsection{Norm of Subgaussian Random Vector}

Let $V = [V_1,...,V_s] \in \R^s$ be a subgaussian random vector, each with subgaussian norm at most $\sigma$.

Then, with probability at least $1-\delta$, $\| V \|_\infty$ is at most $\sigma \sqrt{ \log \frac{s}{\delta}}$

Since $\|V\|_2 \leq \sqrt{s} \| V\|_\infty$, with probability at least $1-\delta$, $\| V \|_2 \leq \sigma \sqrt{\frac{s}{c} \log \frac{sC}{\delta}}$.

\subsection{$l_\infty$ norm bound on $\hat{r}$}

\begin{align*}
\hat{r}_i &= Y_i - \sum_{k \in S} \hat{f}_k (X_k^{(i)}) \\
	&= f^*(X_S^{(i)}) + \epsilon_i - \bar{f^*} - \bar{\epsilon}- \sum_{k \in S} \hat{f}_k(X_k^{(i)}) \\
	&=  f^*(X_S^{(i)}) - \bar{f^*} - \sum_{k \in S} \hat{f}_k(X_k^{(i)}) + \epsilon_i - \bar{\epsilon}
\end{align*}

Where $\bar{f^*} = \frac{1}{n} \sum_{i=1}^n f^*(X_S^{(i)})$ and likewise for $\bar{\epsilon}$.\\

\textbf{Epsilon Bound}
Suppose $\epsilon_i$ is subgaussian with subgaussian norm $\sigma$. For a single $\epsilon_i$, we have that $P( |\epsilon_i| \geq t ) \leq C \exp( - c \frac{1}{\sigma^2} t^2)$. Therefore, with probability at least $1-\delta$, $|\epsilon_i| \leq \sigma\sqrt{ \frac{1}{c} \log \frac{C}{\delta}}$.

By union bound, with probability at least $1-\delta$, $\max_i |\epsilon_i| \leq \sigma \sqrt{ \frac{1}{c} \log \frac{2nC}{\delta}}$.\\

Also, $|\bar{\epsilon}| \leq \sigma \sqrt{\frac{c}{n} \log \frac{C}{\delta}}$ with probability at least $1 - \delta$.\\

\textbf{Bounded Function Bound}

Suppose $|f^*(x)| \leq B$ and $|\hat{f}_k(x_k)| \leq B$ for all $k$.

Then $|\bar{f^*}| \leq B$ as well, and $|f^*(X_S^{(i)}) - \bar{f^*} - \sum_{k \in S} \hat{f}_k(X_k^{(i)})| \leq (s+2) B$ therefore.

\textbf{Lipschitz Function Bound}
Suppose $f^*$ and $\hat{f}_k$ are all Lipschitz with Lipschitz constant at most $L_{\max}$. The composite function $\sum_{k \in S} \hat{f}_k$ is also $L_{\max}$-Lipschitz.

Let $u = \bar{ f^* }$. $X_S \mapsto f^*(X_S) - \bar{f^*}$ is still $L_{\max}$-Lipschitz of course.

Define $h(X^{(i)}_S) \coloneqq f^*(X^{(i)}_S) -u - \sum_{k \in S} \hat{f}_k(X^{(i)}_k)$ as a short-hand. $h$ is therefore Lipschitz with Lipschitz constant at least $2L_{\max}$.

Because $\sum_i h(X^{(i)}_S) = 0$, it must be that for some $i$, $h(X^{(i)}_S) \geq 0$ and for some $i$, $h(X^{(i)}_S) \leq 0$.

\[
\trm{Therefore,}\qquad \max_i |h(X^{(i)}_S)|\leq \max_{i,j} | h(X^{(i)}_S) - h(X^{(j)}_S)|
\]

\begin{align*}
| h(X^{(i)}_S) - h(X^{(j)}_S)|& \leq 2L_{\max} \| X^{(i)}_S - X^{(j)}_S \| \\
&\leq 2L_{\max} (\|X^{(i)}_S\| + \|X^{(j)}_S\|) \\
&\leq 4L_{\max} \max_i \|X^{(i)}_S\|
\end{align*}

Therefore, $\max_i | h(X^{(i)}_S)| \leq 4 L_{\max} \max_i \|X^{(i)}_S\|$.

Using the concentration inequality on norms of subgaussian random vectors and a union bound, we get that, with probability at least $1-\delta$, $\max_i \|X^{(i)}_S\| \leq \sqrt{ \frac{s}{c} \log \frac{nsC}{\delta} }$ (assuming that each coordinate of $X_S$ has subgaussian norm $1$.\\

Taking union bound across deviation for $\epsilon$ and the Lipschitz function term, we have that, with probability at least $1 - \delta$,
\[
\| \hat{r} \|_\infty \leq \sigma L_{\max} \sqrt{ \frac{s}{c} ( \log sn + \log \frac{C}{\delta})}
\]

\subsection{Sampling Without Replacement}

\begin{lemma} (Serfling) 
Let $x_1,..., x_N$ be a finite list, $\bar{x} = \mu$. Let $X_1,...,X_n$ be sampled from $x$ without replacement.

Let $b = \max_i x_i$ and $a = \min_i x_i$. Let $r_n = 1- \frac{n-1}{N}$. Let $S_n = \sum_i X_i$.
Then we have that
\[
P( S_n - n \mu \geq n \epsilon) \leq \exp( - 2 n \epsilon^2 \frac{1}{r_n (b-a)^2})
\]
\end{lemma}

\begin{corollary}
Suppose $\mu = 0$. 
\[
P( \frac{1}{N} S_n \geq \epsilon) \leq \exp( -2 N \epsilon^2 \frac{1}{(b-a)^2})
\]

And, by union bound, we have that
\[
P( | \frac{1}{N} S_n| \geq \epsilon) \leq 2 \exp( -2 N \epsilon^2 \frac{1}{(b-a)^2})
\]

\end{corollary}

A simple restatement. With probability at least $1- \delta$, the deviation $| \frac{1}{N} S_n$ is at most $ (b-a) \sqrt{ \frac{1}{2N} \log \frac{2}{\delta}}$.s

\begin{proof}
\[
P( \frac{1}{N} S_n \geq \epsilon) = P( S_n \geq \frac{N}{n} n \epsilon) \leq \exp( - 2 n \frac{N^2}{n^2} \epsilon^2 \frac{1}{r_n (b-a)^2} ) 
\]

We note that $r_n \leq 1$ always, and $n \leq N$ always. 
\[
\exp( - 2 n \frac{N^2}{n^2} \epsilon^2 \frac{1}{r_n (b-a)^2} )  \leq \exp( - 2 N \epsilon^2 \frac{1}{(b-a)^2})
\]
This completes the proof.

\end{proof}

\subsection{Bounding $\frac{1}{n} \hat{r}^\tran \max( X, X_{(j)} \mathbf{1})$}
\[
\frac{1}{n} \hat{r}^\tran \max(X, X_{(j)} \mathbf{1}) = \frac{1}{n} \sum_{i=1}^n \hat{r}_i \max(X_i, X_{(j)}) = \frac{1}{n} \sum_{i=1}^n \hat{r}_i X_i \delta(ord(i) < j) + \frac{1}{n} X_{(j)} \mathbf{1}_A^\tran \hat{r}_A
\]

Where $A = \{ i : ord(i) \geq j\}$

We will bound both terms.\\

\textbf{Term 1.}
\[
\trm{Want to bound} \qquad F(X_1,...,X_n) \coloneqq \frac{1}{n}\sum_{i=1}^n \hat{r}_i X_i \delta(ord(i) < j) 
\]

First, we assume that $X_i$ is bounded in the range $[-b,b]$.

We claim then that $F$ is coordinatewise-Lipschitz. Let $X = (X_1, X_2,...,X_n)$ and $X' = (X'_1, X_2, ..., X_n)$ differ only on the first coordinate. \\

The order of coordinate $i$ in $X$ and $X'$ can change by at most $1$ for $i \neq 1$. Therefore, of the $j-1$ terms of the series, at most 2 terms differ from $F(X)$ to $F(X')$. Therefore,

$$ | F(X_1, ..., X_n) - F(X'_1, ..., X_n) | \leq \frac{4 b \|\hat{r}\|_\infty}{n} $$

By McDiarmid's inequality therefore, 
\[
P( | F(X) - \E F(X) | \geq t ) \leq C \exp( - c n\frac{t^2}{ (4 b \|\hat{r}\|_\infty)^2})
\]

By symmetry and the fact that $\hat{r}$ is centered, $\E F(X) = 0$.

We can fold the 4 into the constant $c$. With probability $1-\delta$, $|F(X)| \leq b \|\hat{r}\|_\infty \sqrt{   \frac{1}{cn} \log \frac{C}{\delta}}$. \\

\textbf{Term 2:}

\[
\trm{Want to bound}   \frac{1}{n} X_{(j)} \mathbf{1}_A^\tran \hat{r}_A
\]

$A$ is a random set and is probabilistically independent of $\hat{r}$.  $\mathbf{1}_A^\tran \hat{r}_A$ is the sum of a sample of $\hat{r}$ without replacement. Therefore, according to Serfling's theorem, with probability at least $1-\delta$, $|\frac{1}{n} \mathbf{1}_A^\tran \hat{r}_A|$ is at most $\| \hat{r} \|_\infty \sqrt{ \frac{1}{cn} \log \frac{C}{\delta}}$.

Since $|X_{(j)}|$ is at most $b$, we obtain that with probability at least $1-\delta$, $| \frac{1}{n} X_{(j)} \mathbf{1}_A^\tran \hat{r}_A| \leq b\| \hat{r} \|_\infty \sqrt{ \frac{1}{cn} \log \frac{C}{\delta}}$.\\

\textbf{All Together}

Taking union bound across $p$ and $n$, we have that with probability at least $1-\delta$,

\[
| \frac{1}{n} \max(X, X_{(j)}\mathbf{1})^\tran \hat{r}| \leq b \|\hat{r}\|_\infty \sqrt{ \frac{1}{c} \frac{1}{n} \log \frac{np C }{\delta}}
\]

Taking union bound and substituting in the probabilistic bound on $\|\hat{r}\|_\infty$, we get that with probability at least $1-\delta$,

$| \frac{1}{n} \max(X, X_{(j)}\mathbf{1})^\tran \hat{r}|$ is at most 
\[
\sigma L_{\max} b \sqrt{ \frac{1}{c} \frac{s}{n} ( \log (spn) + \log \frac{C}{\delta})^2}
\]

\newpage

\textbf{DEPRECATED!}\\ %===%

\section{Additive Model}
\subsection{Proof Sketch}

FACT: A bounded convex function is Lipschitz. \\

FACT: If $f_k(x_k)$ are $L$-Lipschitz, then $\sum_k f_k(x_k)$ is also $L$-Lipschitz.

\[
| \sum_k f_k(x_k) - \sum_k f_k(x'_k) |^2 \leq \sum_k |f_k(x_k) - f_k(x'_k)|^2 \leq \sum_k L^2| x_k - x'_k |^2 \leq L^2 \| x - x' \|_2^2
\]

FACT: if $X$ is sub-gaussian random vector, then $g(X)$ is also sub-gaussian if $g$ is Lipschitz. This is true by McDiarmid's inequality.\\

ASSUMPTION: $X_k$ for $k=1,...,p$ are independent. We can use weaker assumptions. \\

Let $\hat{f}_k$ be the optimal solution to the optimization
\[
\frac{1}{2n} \| Y - \sum_{s=1}^p \hat{f}_s(X_s) \|_2^2 + \lambda\sum_{s=1}^p \| \partial \hat{f}_s(X_s)\|_2 \qquad \hat{f}_s \trm{ convex and } \sum_i \hat{f}_s (x_{is}) = 0 .
\]

ASSUMPTION: $Y$ has mean zero, $\hat{f}_k$ for all $k$ has mean zero, and $\hat{f}_k$ is Lipschitz with an universal upper bound on the Lipschitz constant.\\

CLAIM: If we set $\lambda = O \sqrt{\frac{\log (np) }{ n }}$, then $\hat{f}_k$ for an irrelevant $k$ is zero.

We fix an irrelevant dimension $k$, and define the residue $\hat{y} = y - \sum_{k' \neq k} \hat{f}_{k'}(X_{k'})$.

Using the reformulation, we can re-write the optimization, in term of just $f_k$, as the following.

\[
\min_{d, c}  \frac{1}{2n} \| \hat{y} - \Delta d - c \|_2^2 + \lambda \sum_{i=1}^n d_i + \lambda|d_0|  \qquad \trm{s.t.} \forall i>0, \, d_i \geq 0
\]
Where $d$ is a $n$-dim vector, $c$ scalar is the centering variable, and $\Delta$ is the $n \times n$ matrix of inter-sample distances.\\

Taking the subgradient with respect to $d$, we have that
\[
\frac{1}{n} (-\hat{y}^\tran \Delta + d^\tran \Delta^\tran \Delta ) + \lambda \mathbf{u} - \mu = 0
\]

where $\mathbf{u}$ is a $n$-dim vector whose first coordinate is $\partial | d_0 |$ and whose all other coordinates are $1$. $\mu$ is a $n$-dim vector whose first coordinate is $0$ and all other coordinates are the Lagrangian multipliers.\\

We now want to verify that if we substitute $d = 0$ into the subgradient, the subgradient equals 0. If $d=0$, then we can substitute $\mathbf{u} = \mathbf{1}$.

All we need to verify then is that $\lambda \mathbf{1} \geq \frac{1}{n} \hat{y}^\tran \Delta$.\\

Note that $\hat{y}$ is a sub-gaussian random vector with mean zero. Because $\hat{y}$ and $\Delta$ are independent, $\hat{y}^\tran \Delta$ is also a sub-gaussian mean-zero random vector. Note also that $\Delta$ is entirely positive.\\

The empirical mean $ \frac{1}{n} \hat{y}^\tran \Delta$ is then on the order of $\sqrt{ \frac{1}{n}}$ with high probability.

We need to take union bound across the $n$-dimensions of $\hat{y}^\tran \Delta$ and across the $p$ dimensions.

Therefore, setting $\lambda = O \sqrt{\frac{\log (np)}{n}}$ suffices.

\subsection{Reformulation}

The objective is, again
\[
\frac{1}{2n} \| Y - \sum_{s=1}^p \hat{f}_s(X_s) \|_2^2 + \sum_{s=1}^p \| \partial \hat{f}_s(X_s)\|_2 \qquad \hat{f}_s \trm{ convex and } \sum_i \hat{f}_s (x_{is}) = 0
\]

Let us try to simplify this form. We can replace the identifiability constraint with $\hat{f}_s( 0 ) = 0$ or $\hat{f}_s(-1) = 0$. Let's go with the latter formulation and define $x_{s(0)}$ as $-1$.

Let $x_{s(1)} ,..., x_{s(n)}$ be the $n$ samples arranged from small to large.\\

Then  $\hat{f}_s(x_{s(1)}) = \hat{f}_s(x_{s(0)}) + \hat{\beta}_{s(0)} (x_{s(1)}-x_{s(0)} )$

and $\hat{f}_s(x_{s(2)}) = \hat{\beta}_{s(0)} (x_{s(1)}-x_{s(0)} ) + \hat{\beta}_{s(1)} (x_{s(2)}-x_{s(1)} )$\\

Some additional transformation. Let's define $\hat{d}_{s(0)}$ as the gradient increments. $\hat{d}_{s(0)} = \hat{\beta}_{s(0)}$, and $\hat{d}_{s(1)} = \hat{\beta}_{s(1)} - \hat{\beta}_{s(0)}$.\\

Therefore, $\hat{\beta}_{s(i)} = \sum_{j \leq i} \hat{d}_{s(i)}$.

\[
\hat{f}_s(x_{s(1)}) = \hat{f}_s(x_{s(0)}) + \hat{d}_{s(0)} ( x_{s(1)} - x_{s(0)})
\]

\begin{align*}
\hat{f}_s(x_{s(2)}) &= \hat{f}_s(x_{s(1)}) + \hat{\beta}_{s(1)} ( x_{s(2)} - x_{s(1)}) \\
 &= \hat{d}_{s(0)} ( x_{s(1)} - x_{s(0)}) + (\hat{d}_{s(0)}+\hat{d}_{s(1)}) (x_{s(2)} - x_{s(1)}) \\
 &= \hat{d}_{s(0)} (x_{s(2)}-x_{s(0)} ) + \hat{d}_{s(1)} (x_{s(2)}-x_{s(1)} )
\end{align*}

\[
\hat{f}_s(x_{s(i)}) = \hat{d}_{s(0)} ( x_{s(i)} - x_{s(0)}) + \hat{d}_{s(1)} ( x_{s(i)} - x_{s(1)}) + ... + \hat{d}_{s(i-1)} ( x_{s(i)} - x_{s(i-1)})
\]

Define $\Delta(j, x_{si}) = 0$ if $\trm{order}(i) \leq j$, $x_{si} - x_{s(j)}$ else. With this definition, we can re-write

\[
\hat{f}_s(x_{si}) = \hat{d}_{s}^\tran \Delta(x_{si})
\]

With the simple constraint that all $\hat{d}_{si} \geq 0$ for $i > 0$.\\

\textbf{Notation:} We will also define $\Delta_s$ as a $(n \times n)$ matrix whose $i,j$-th entry is $\Delta(j, x_{si})$.

With this short-hand, we can write $\hat{f}_s(x_s) = \Delta_s\hat{d}_s$.

\subsection{Consistency with Short-hand}

We can now attempt the consistency proof with our short-hand notation.

First, what is the objective? $\| \partial \hat{f}_s(X_s) \|_\infty$ becomes $\| \hat{d}_s \|_1$.

The objective is therefore
\[
\frac{1}{2n} \| Y - \sum_{s=1}^p \Delta_s \hat{d}_s \|_2^2 + \sum_{s=1}^p \|\hat{d}_s \|_1 .
\]

Taking the same form as before, we have
\[
\frac{1}{2n} \left\{ 
  \| \sum_s \Delta_s ( \hat{d}_s - d^*_s) \|_2^2 +
\sum_s \langle \Delta_s ( \hat{d}_s - d^*_s) , \epsilon \rangle \right\} \leq
\lambda \sum_s \|d^*_s \|_1 - \lambda \sum_s \|\hat{d}_s \|_1
\]

The cross term can be bounded.
\begin{align*}
\frac{1}{2n} |\sum_s \langle \Delta_s ( \hat{d}_s - d^*_s), \epsilon \rangle| &\leq  \frac{1}{2} \sum_s \langle \hat{d}_s - d^*_s, \frac{1}{n}\Delta_s^\tran \epsilon \rangle \\
&\leq \frac{1}{2} \sum_s \| \hat{d}_s - d_s^* \|_1 \| \frac{1}{n}\Delta_s^\tran \epsilon \|_\infty \\
&\leq \frac{1}{2} \lambda \sum_s \| \hat{d}_s - d_s^* \|_1
\end{align*}

Continuing, we have
\[
\frac{1}{2n}\| \sum_s \Delta_s ( \hat{d}_s - d^*_s) \|_2^2 + \frac{\lambda}{2} \sum_{s \in S^c} \| \hat{d}_s \|_1 \leq \frac{3\lambda}{2} \sum_{s \in S} \| \hat{d}_s - d^*_s \|_1  
\]

Inside the loss function, we have $\Delta v$ where $\Delta$ is a $n \times n \times p$ Tensor and $v = \hat{d} - d^*$ is a $p \times n$ matrix.

The tensor multiplication can be re-written 
\[
\sum_{s,i} \Delta_{si} v_{si} = \tilde{\Delta} \tilde{v} \qquad \trm{where $\tilde{\Delta}$ is $n \times np$, $\tilde{v}$ is $np$-vector.}
\]

\subsection{Consistency}

First we have to establish framework and notation. The optimization can be written in many different ways, with or without constraint. 
\[
\frac{1}{2n} \| Y - \sum_{s=1}^p \hat{f}_s(X_s) \|_2^2 + \sum_{s=1}^p \| \partial \hat{f}_s(X_s)\|_2 \qquad \hat{f}_s \trm{ convex}
\]
$X \in \R^{n\times p}$ and $X_s \in \R^n$ where $s = 1,...,p$. $\hat{f}_s(X_s) = [ \hat{f}_s(X_{is}) ]_i$ is also a $n$-dim vector.

\textbf{Step 1:}\\

Transforming the objective. We claim that each $\hat{f}_s$ is associated with pairs of real numbers $\{ \hat{\alpha}_{s k}, \hat{\beta}_{s k} \}_k$ such that $\hat{f}_s(X_{ks}) = \hat{\alpha}_{s k} + \hat{\beta}_{s k} X_{ks}$.

The alphas and betas must satisfy $\hat{\alpha}_{s k'}, \hat{\beta}_{s k'} = \argmax_k \hat{\alpha}_{s k} + \hat{\beta}_{s k} X_{k' s}$.\\

Our starting point will be the same as before. 
\begin{align*}
& Y = \sum_{s=1}^p f^*_s(X_s) + \epsilon \\
& \frac{1}{2n} \| Y - \sum_{s=1}^p \hat{f}_s(X_s) \|_2^2 + \sum_{s=1}^p \| \partial \hat{f}_s(X_s)\|_{2,1} \geq \frac{1}{2n} \| Y - \sum_{s=1}^p f^*_s(X_s) \|_2^2 + \sum_{s=1}^p \| \partial f^*_s(X_s)\|_2
\end{align*}

\textbf{Note to self:} $f^*_s$'s are not unique. Will this be a problem?\\

We can do algebra and get the following conclusion.

\[
\frac{1}{2n} \left\{ \| \sum_s f^*_s(X_s) - \sum_s \hat{f}_s(X_s) \|_2^2 + \sum_s \langle f^*_s(X_s) - \hat{f}_s(X_s), \epsilon \rangle \right\} \leq \lambda \sum_s \| \partial f^*_s(X_s)\|_2 - \lambda  \sum_s \| \partial \hat{f}_s(X_s)\|_2 
\]

\textbf{Cross-term.} We fix $s$. 
\begin{align*}
& \langle f^*_s(x_s) - \hat{f}_s(x_s), \epsilon \rangle \\
=& \sum_{k=1}^n [ \alpha^*_{sk} + \beta^*_{sk} x_{ks} - \hat{\alpha}_{sk} - \hat{\beta}_{sk} x_{ks} ] \epsilon_k \\
=& \sum_{k=1}^n [\alpha^*_{sk} - \hat{\alpha}_{sk}] \epsilon_k + [\beta^*_{sk} - \hat{\beta}_{sk}] \epsilon_k x_{ks} \\
\leq & \| \alpha^*_s - \hat{\alpha}_s \|_2 \| \epsilon \|_2 + \| \beta^*_s - \hat{\beta}_s \|_2 \| x_s \epsilon \|_2 \quad \trm{ $x_s \epsilon$ is a $n$-dim vector}
\end{align*}

It follows that $\| \epsilon \|_2$ and $\| x_s \epsilon \|_2$ are both on the order of $O(\sqrt{n})$.\\

Here lies the difficulty. In vanilla Lasso, there is only ``one'' parameter vector, $\beta$. Therefore, the inner product is easy to bound:

\begin{align*}
\frac{1}{n} \langle X \Delta, \epsilon \rangle = \langle \Delta, \frac{1}{n}X^\tran \epsilon \rangle
\end{align*}

And $\frac{1}{ n} X^\tran \epsilon$ is the empirical average of samples from a mean-zero distribution.

In new Lasso, the problem is that there are ``as many'' parameter vectors as there are samples. Let $i$ index a sample and let $s$ index a dimension.

\begin{align*}
& \frac{1}{n} \sum_i X_i^\tran \Delta_i \epsilon_i \\
=& \frac{1}{n} \sum_i \sum_s X_{is} \Delta_{is} \epsilon_i \\
=& \frac{1}{n} \sum_s \sum_i \Delta_{is} ( X_{is} \epsilon_i )
\end{align*}

The problem now is that we are forced to take the norm of $( \Delta_{is} X_{is} )_i$, which is of too large of an order.

\textbf{Continuing.}

First, we have to formally define $\| \partial f_s^*(X_s) \|_2$, which we define as $\| \alpha^*_s \|_2 + \| \beta^*_s\|_2$.

Therefore, by setting $\lambda$ appropriately, the derivation becomes:

\begin{align*}
 \frac{1}{2n} \| \sum_s f^*_s(X_s) - \sum_s \hat{f}_s(X_s) \|_2^2 - \frac{\lambda}{2} \| f^* - \hat{f} \|_{2,1}  &\leq \lambda \|f^*\|_{2,1} - \lambda  \| \hat{f} \|_{2,1}  \quad \trm{note $\|f^* \|_{2,1} \equiv \sum_s \| f^*_s \|_2$.}\\
\frac{1}{2n} \| \sum_s f^*_s(X_s) - \sum_s \hat{f}_s(X_s) \|_2^2 &\leq \frac{3\lambda}{2} \| f^* - \hat{f}_S \|_{2,1} - \frac{\lambda}{2} \| \hat{f}_{S^c} \|_{2,1}
\end{align*}

\section{Simultaneous Update for $h$ and $S$}

\subsection{Stopping Criteria}

If we can condense the updates, then it is possible to have a sensible stopping criteria for ADMM.

The stopping criteria check the fulfillment of the stationarity condition and the primal feasibility condition.

\subsection{The Derivations}

The matrix of the form $\sum_{ij} (\mathbf{e}_j - \mathbf{e}_i) (\mathbf{e}_j - \mathbf{e}_i)^\tran$ has an explicit form:

It is $2n -2$ on the diagonal and $-2$ everywhere else.\\

We do this for the vanilla implementation--no quadratic term, no additive asssumption etc.

We must solve the following linear equation:
\begin{align*}
( \frac{1}{\mu} \mathbf{I} + 2L ) \mathbf{h} &= \left( \frac{1}{\mu} \mathbf{y} + \sum_{ij} (\mathbf{e}_j - \mathbf{e}_i) ((x_j - x_i)^\tran B_i + S_{ij} - \frac{1}{\mu} W_{ji}) \right) \\
S_{ji} &= h_j - h_i - B_i^\tran (x_j - x_i) + \frac{1}{\mu} W_{ji}
\end{align*}

$L$ is a matrix of the form $D - A$ where $A$ is the all ones matrix and $D$ is the degree matrix of $A$. Therefore, $L$ is all negative ones everywhere except for the diagonal, where it is $n-1$.

\textbf{Step 1.} We separate the variables.
\begin{align*}
( \frac{1}{\mu} \mathbf{I} + 2L ) \mathbf{h} - \sum_{ij} (\mathbf{e}_j - \mathbf{e}_i) S_{ij} &= \left( \frac{1}{\mu} \mathbf{y} + \sum_{ij} (\mathbf{e}_j - \mathbf{e}_i) ((x_j - x_i)^\tran B_i - \frac{1}{\mu} W_{ji}) \right) \\
S_{ji} - h_j + h_i &= - B_i^\tran (x_j - x_i) + \frac{1}{\mu} W_{ji}
\end{align*}

\textbf{Step 2.} We describe the matrix form.
\[
\boxed{\trm{Big Matrix}} \left[\begin{array}{c} \mathbf{h} \\ \mathbf{S} \end{array}\right] = \boxed{\trm{Big Vector}}
\]

The Big Matrix is $(n+n^2) \times (n+n^2)$ and is block structured:
\[
\left[\begin{array}{cc} 
\mu^{-1} \mathbf{I} + 2L & \trm{Operator h-S} \\
\trm{Operator S-h} & \mathbf{I}_{|S|}
\end{array}\right]
\]

Operator h-S is $n \times n^2$ and represented by $-\sum_{ij} (\mathbf{e}_j - \mathbf{e}_i) S_{ji}$, which is the output vector when we multiply by $S$

Operator S-h is $n^2 \times n$ and represented by $ - h_j + h_i$, which is the $(j,i)$-th entry of the output $n^2$-vector when we multiply by $h$.

We must also describe what the \fbox{Big Vector} is. Its form can be directly read from the equation array in step 1.

\textbf{Step 3.} We apply schur complement to invert.

Schur complement, abstractly written, is the following identity:
\[
\left[ \begin{array}{cc} A & B \\ C & D \end{array} \right]^{-1} = 
\left[ \begin{array}{cc} I_p & 0 \\ -D^{-1} C & I_q \end{array} \right]
\left[ \begin{array}{cc} (A - BD^{-1}C)^{-1} & 0 \\ 0 & D^{-1} \end{array} \right]
\left[ \begin{array}{cc} I_p & - B D^{-1} \\ 0 & I_q \end{array} \right]
\]

In our situation, $D$ is identity.

\textbf{Step 4.} We perform step 1 of the Schur complement

We now apply a $n+n^2 \times n+n^2$ matrix onto a $n + n^2$-vector and describe the resulting vector. We do not want to store a large matrix in memory.

Let $\left[ \begin{array}{c} \mathbf{h}' \\ \mathbf{S}' \end{array} \right]$ be the input vector.

The result is
\[
\left[ \begin{array}{c}
\mathbf{h}' + \sum_{ij} (\mathbf{e}_j - \mathbf{e}_i) \mathbf{S}'_{ji} \\
\mathbf{S}'
\end{array}\right]
\]

\textbf{Step 5.} We perform step 2 of the Schur complement

We must first compute the composition Operator h-S $\circ$ Operator S-h.

Let $\mathbf{h}'$ be a vector, then after Operator S-h, we have a matrix whose $(j,i)$ entry is $ - \mathbf{h}'_j + \mathbf{h}'_i$.

After an application of Operator h-S, we now have a vector of the form $- \sum_{ij} (\mathbf{e}_j - \mathbf{e}_i) ( - \mathbf{h}'_j + \mathbf{h}'_i)$.

That can be re-written as: $\sum_{ij} (\mathbf{e}_j - \mathbf{e}_i) (\mathbf{e}_j - \mathbf{e}_i)^\tran \mathbf{h}'$.

Hence, the combined operator is $2L$.

Because $A$ is $\mu^{-1} \mathbf{I} + L$, the second step is just a multiplication of the first part by $\mu$ and an identity operation on the second part.

Thus, continuing our calculation from the previous step, the vector $\mathbf{h}'$ and $\mathbf{S}'$ have now become:

\[
\left[ \begin{array}{c}
\mu \mathbf{h}' + \mu \sum_{ij} (\mathbf{e}_j - \mathbf{e}_i) \mathbf{S}'_{ji} \\
\mathbf{S}'
\end{array}\right]
\]

\textbf{Step 6.} We perform step 3 of the Schur complement

We must now apply negative Operator S-h onto $\mu \mathbf{h}' + \mu \sum_{ij} (\mathbf{e}_j - \mathbf{e}_i) \mathbf{S}'_{ji} $

Applying Operator S-h onto $\mathbf{h}'$ results in $[- \mathbf{h}'_s + \mathbf{h}'_t]_{s,t}$

The second part is harder. A naive application gives $[\sum_{ij} (\mathbf{e}_j(s) - \mathbf{e}_i(s) - (\mathbf{e}_j(t) - \mathbf{e}_i(t))) \mathbf{S}'_{ji} ]_{s,t}$\\

We may distribute the $S'_{ji}$ into the differences and get:
\[
\sum_i S'_{si} - \sum_j S'_{js} - \sum_i S'_{ti} + \sum_j S'_{jt}
\]
To check for correctness, we remember that the constraints are, in order, $j=s$, $i=s$, $j=t$, $i=t$.

\textbf{Implementation Note:} We have: rowsum(s)-colsum(s)-rowsum(t)+colsum(t). For efficient implementation, we first compute all row and column sums of $S'$. We then construct a matrix whose $(s,t)$-th entry is $rowsum(s) - rowsum(t)$.

\section{ADMM with Gap and Quadratic Term}

Let $B$ be a $p\times n$ matrix, and let $D$ also be a $p\times n$ matrix. $C$ is then a $p\times 2n$ matrix; let $C = [C_B, C_D]$.

\begin{align*}
\min_{\mathbf{h}, \mathbf{B}, \mathbf{C},\mathbf{S},\mathbf{D}}
& \frac{1}{2n}\sum_{i=1}^n(y_i - h_i)^2 + \lambda \|C\|_{\infty,1} + \gamma \frac{1}{n} \sum_{ij} S_{ji} \\
\trm{s.t.} & C_B = B, C_D = D, S_{ji} \geq 0 \\
	& S_{ji} = h_j - h_i - B_i^\tran(X^j - X^i) - \sum_{k=1}^p D_{ik} (X^j_k - X^i_k)^2
\end{align*}

The corresponding ADMM objective is similar to before:
\begin{align*}
\min_{\mathbf{h},\mathbf{B},\mathbf{C},\mathbf{D},\mathbf{S},\mathbf{W},\mathbf{M}} &
\frac{1}{2}\sum_{i=1}^n (y_i - h_i)^2 + \lambda \|C\|_{\infty,1} + \gamma \sum_{ij}S_{ji} \\
& +\sum_{ij} (W_{ji} \cdot [ h_j - h_i - B_i^\tran(X^j - X^i) - \sum_{k=1}^p \mathbf{D}_{ik} (X^j_k - X^i_k)^2 - S_{ji}]) \\
& + \frac{\mu}{2} \sum_{ij} ( h_j - h_i - B_i^\tran (X^j - X^i) -\sum_{k=1}^p \mathbf{D}_{ik} (X^j_k - X^i_k)^2-S_{ji})^2 \\
& + \trace(M^\tran (C - [B,D])) + \frac{\mu}{2} \| C - [B,D]\|^2 \trm{ s.t. } S_{ji} \geq 0
\end{align*}

I need to introduce additional columns to $C, M$, introduce $D$. 
The updates for $\mathbf{h}$ and $\mathbf{B}$ and $\mathbf{W}$ are easy. The updates for $\mathbf{C},\mathbf{W},\mathbf{M}$ are easy.\\

\subsection{Update for $\mathbf{C}$}

The original update was:
\begin{align*}
& \min_{\mathbf{C}} \lambda \|\mathbf{C}\|_{\infty,1} + \frac{\mu}{2} \| \mathbf{C} - (B - \frac{1}{\mu} M ) \|_2^2 \\
\end{align*}

$\mathbf{C}$ is thus $\mathbf{C}$ subtracts a projection of $B - \frac{1}{\mu} M$ onto $\|\cdot\|_1$ ball of radius $\frac{\lambda}{\mu}$.

The new update is 
\begin{align*}
& \min_{\mathbf{C}} \lambda \|\mathbf{C}\|_{\infty,1} + \frac{\mu}{2} \| \mathbf{C} - ([B,D] - \frac{1}{\mu} M ) \|_2^2 \\
\end{align*}

Same basic code. No change.

\begin{framed}
We take a brief digression to discuss how to solve problems of the form:
\[
\min_{z \in \R^M} \| z - x\|_2^2 + \nu \|z \|_q
\]

The subgradient optimality condition states that 
\[
x - z \in \frac{\nu}{2} \partial \| z \|_q
\]

Define $\Psi^*(y) = \sup_z \langle z, y \rangle - \Psi(z)$ as the Fenchel conjugate of $\Psi(z) = \frac{\nu}{2} \| z \|_q$.

FACT: For any Fenchel conjugate pairs $f,g$, it holds that $y \in \partial f(x)$ iff $x \in \partial g(y)$.

Therefore, we have that 
\[
z \in \partial \Psi^*(x - z) \Rightarrow x \in x-z - \partial \Psi^*(x-z)
\]

It is claimed that $\Psi^*(y) = \Xi_{B^{q'}(\nu/2)}(y)$ is the indicator function. Why is this?
Let us look at $\sup_x \langle x, y\rangle - \Psi(x)$ for different values of $y$. If $\|y\|_{q'} \leq \frac{\nu}{2}$, then $\frac{\nu}{2} \|x\|_q \geq \langle x,y\rangle$ for all $x$ because $\|y\|_{q'} = \sup_x \langle \frac{x}{\|x\|_q},y\rangle $. 
If $\|y\|_{q'} > \frac{\nu}{2}$, then there would exist an $x$ such that $\langle x, y\rangle > \frac{\nu}{2} \|x\|_q$.

FACT: $y \in w + \partial \Xi_C(w)$ for an $\infty$-indicator function $\Xi_C(w)$ of a convex set $C$ iff we have that $w = \argmin_{w' \in C} \|w' - y\|_2^2$. 
PROOF:

Suppose the primus holds. Then $(y-w)^\tran (w' - w) \leq \Xi_C(w') - \Xi_C(w)$ for all $w' \in C$, this implies that $(y-w)^\tran (w' - w) \leq 0$ for all $w' \in C$.

The optimization can be restated with the indicator function: $\min_{w'} \|w' - y \|_2^2 \trm{ s.t. } \Xi_C(w') \leq 0$. We can then form the Lagrangian.

Note: $v \in \partial \Xi_C(w)$ is equivalent to $v$ in the tangent cone of $C$ at $w$.

In any case, we establish the connection. $x-z$ is the projection of $x$ onto the $\frac{\nu}{2}$ ball of the dual norm.

\end{framed}

\subsection{Update for $\mathbf{B}$}

Recall $\mathbf{B}$ is $p\times n$. The original update. $B_i$ is column of $\mathbf{B}$
\[
\min_{B_i} \sum_{j=1}^n \frac{\mu}{2} (h_j - (h_i + B_i^\tran (X^j - X^i) + S_{ji} ) + \frac{1}{\mu}W_{ji})^2 + \frac{\mu}{2} \| B_i - (\mathbf{C}_i + \frac{1}{\mu} \mathbf{M}_i) \|_2^2
\]
\[
B_i = \left( \mathbf{I} + \sum_{j=1}^n (X^j - X^i)(X^j - X^i)^\tran \right)^{-1} 
	\left( \mathbf{C}_i + \frac{1}{\mu}\mathbf{M}_i + \sum_{j=1}^n (X^j - X^i)(h_j - h_i - S_{ji} + \frac{1}{\mu} W_{ji}) \right)
\]

The new forms should be:
\[
\min_{B_i} \sum_{j=1}^n \frac{\mu}{2} \left(h_j - (h_i + B_i^\tran (X^j - X^i) + \sum_{k=1}^p \mathbf{D}^i_k (X^j_k - X^i_k)^2 + S_{ji} ) + \frac{1}{\mu}W_{ji}\right)^2 + \frac{\mu}{2} \| B_i - (\mathbf{C}_{B, i} + \frac{1}{\mu} \mathbf{M}_{B,i}) \|_2^2
\]
\[
B_i = \left( \mathbf{I} + \sum_{j=1}^n (X^j - X^i)(X^j - X^i)^\tran \right)^{-1} 
	\left( \mathbf{C}_{B,i} + \frac{1}{\mu}\mathbf{M}_{B,i} + \sum_{j=1}^n (X^j - X^i)(h_j - h_i - \sum_{k=1}^p \mathbf{D}^i_k (X^j_k - X^i_k)^2+ S_{ji} + \frac{1}{\mu} W_{ji}) \right)
\]

\subsection{Updates for $\mathbf{h}$}

Before
\[
\min_{\mathbf{h}} \frac{1}{2} \sum_{i=1}^n (y_i - h_i)^2 + \sum_{j=1}^n \frac{\mu}{2} \left(h_j - (h_i + B_i^\tran (X^j - X^i) + S_{ji} ) + \frac{1}{\mu}W_{ji}\right)^2
\]
\begin{align*}
h = & \left( \frac{1}{\mu} \mathbf{I} + \sum_{ij} (e_j - e_i)(e_j - e_i)^\tran\right)^{-1} \\
	& \left( \frac{1}{\mu} y + \sum_{ij} (e_j - e_i)((X^j - X^i)^\tran B_i + S_{ji} - \frac{1}{\mu} W_{ji} )\right)
\end{align*}

After
\[
\min_{\mathbf{h}} \frac{1}{2} \sum_{i=1}^n (y_i - h_i)^2 + \sum_{j=1}^n \frac{\mu}{2} \left(h_j - (h_i + B_i^\tran (X^j - X^i) + \sum_{k=1}^p \mathbf{D}^i_k (X^j_k - X^i_k)^2 + S_{ji} ) + \frac{1}{\mu}W_{ji}\right)^2
\]
\begin{align*}
h = & \left( \frac{1}{\mu} \mathbf{I} + \sum_{ij} (e_j - e_i)(e_j - e_i)^\tran\right)^{-1} \\
	& \left( \frac{1}{\mu} y + \sum_{ij} (e_j - e_i)((X^j - X^i)^\tran B_i + \sum_{k=1}^p \mathbf{D}^i_k (X^j_k - X^i_k)^2 + S_{ji} - \frac{1}{\mu} W_{ji} )\right)
\end{align*}

Note: $\sum_{j} e_j M_{ij}$ is the $i$-th row of $M$.

\subsection{Update for $\mathbf{S}$}

Before
\[
\min_{S_{ji}} \frac{\mu}{2} \left(h_j - (h_i + B_i^\tran (X^j - X^i)  + S_{ji} ) + \frac{1}{\mu}W_{ji}\right)^2  \trm{ s.t. } S_{ji} \geq 0
\]
\[
S_{ji} = \max(h_j - (h_i + B_i^\tran (X^j - X^i)) + \frac{1}{\mu} W_{ji}, 0)
\]
After
\[
\min_{S_{ji}} \frac{\mu}{2} \left(h_j - (h_i + B_i^\tran (X^j - X^i) + \sum_{k=1}^p \mathbf{D}^i_k (X^j_k - X^i_k)^2 + S_{ji} ) + \frac{1}{\mu}W_{ji}\right)^2 + \gamma S_{ji} \trm{ s.t. } S_{ji} \geq 0
\]
\[
S_{ji} = \max(h_j - (h_i + B_i^\tran (X^j - X^i) + \sum_{k=1}^p \mathbf{D}^i_k (X^j_k - X^i_k)^2 ) + \frac{1}{\mu} W_{ji} - \gamma, 0)
\]

\subsection{Updates for $\mathbf{W}$, $\mathbf{M}$}

Before:
\[
W_{ji} = W_{ji} + \mu ( h_j - (h_i + B_i^\tran (X^j - X^i) + S_{ji})),  \,\,\, M = M + \mu ( \mathbf{C} - [\mathbf{B},\mathbf{D}])
\]

After:
\begin{align*}
W_{ji} &= W_{ji} + \mu ( h_j - (h_i + B_i^\tran (X^j - X^i) + \sum_{k=1}^p \mathbf{D}^i_k (X^j_k - X^i_k)^2 + S_{ji})) \\
M &= M + \mu ( \mathbf{C} - [\mathbf{B},\mathbf{D}])
\end{align*}

\subsection{Updates for $\mathbf{D}$}

\begin{align*}
\min_D & \sum_{ij} (W_{ji} \cdot [ h_j - h_i - B_i^\tran(X^j - X^i) - \sum_{k=1}^p \mathbf{D}_{ik} (X^j_k - X^i_k)^2 - S_{ji}]) \\
& + \frac{\mu}{2} \sum_{ij} ( h_j - h_i - B_i^\tran (X^j - X^i) -\sum_{k=1}^p \mathbf{D}_{ik} (X^j_k - X^i_k)^2-S_{ji})^2 \\
& + \trace(M_D^\tran (C_D - D)) + \frac{\mu}{2} \| C_D - D\|^2
\end{align*}

Simplifying things, we have 
\[
\min_{D_i} \sum_{j=1}^n \left( h_j - h_i - B_i^\tran (X^j - X^i) - S_{ji} - D_i^\tran \Delta_{ij} + \frac{1}{\mu} W_{ji} \right)^2 \frac{\mu}{2} + \frac{\mu}{2} \| D_i - (C_{D,i} + \frac{1}{\mu} M_{D,i}) \|_2^2
\]
where $\Delta_{ij}$ is a $p$-dim vector and $\Delta_{ij}(k) = (X^j_k - X^i_k)^2$.

Taking a cue from the update for $B_i$, I would guess that the update for $D_i$ is the following:

\begin{align*}
D_i &= \left( \mathbf{I} + \sum_{j=1}^n \Delta_{ij} \Delta_{ij}^\tran \right)^{-1} \\
	& \left( \mathbf{C}_{D,i} + \frac{1}{\mu} \mathbf{M}_{D,i} + \sum_{j=1}^n \Delta_{ij} ( h_j - h_i - S_{ji} - B_i^\tran (X^j - X_i) + \frac{1}{\mu} W_{ji} ) \right)
\end{align*}

\section{High-Dimensional Consistency Analysis}
This proof strategy first bounds denoising error and then use the fact that convex functions are Lipschitz on a compact set to bound the interpolation error.\\

We augment $X$, a $p \times n$ regressor matrix, with one more row of all ones. Then any convex function can be expressed as $\max B^\tran X$ for some $B$ matrix which is $(p+1) \times n$, the max-operation is taken column-wise, that is, over the $n$ columns of $B$ matrix.

Let $\hat{B}$ be the empirical risk minimizer.\\

\textbf{Note:} Even without noise, there could be infinite number of convex functions that agrees with the ground truth on the $n$ data points we sample. The equivalence class of such convex functions become smaller as $n$ increases. We will compare our piece-wise linear convex function against an \emph{arbitrary} piece-wise linear function, which we represent by $B^*$, that agrees with the true function $g^*$ on $X_1,...,X_n$.

 \begin{align*}
\frac{1}{2n} \| Y - \max \hat{B}^\tran X \|_2^2 + \lambda \|\hat{B}\|_{2,1} &\leq
				\frac{1}{2n}\| Y - \max B^{*\tran} X\|_2^2 + \lambda \| B^*\|_{2,1}
\end{align*}

FACT 1: $\frac{1}{2n}\|Y - \max B^{*\tran} X\|_2^2 = \frac{1}{2n}\| \epsilon \|_2^2$.

FACT 2: 
\begin{align*}
\frac{1}{2n} \| Y - \max \hat{B}^\tran X \|_2^2  &= \frac{1}{2n} \| \max B^{* \tran} X - \max \hat{B}^\tran X + \epsilon  \|_2^2\\
&=  \frac{1}{2n} \left\{ \| \max B^{* \tran} X - \max \hat{B}^\tran X  \|_2^2 + \|\epsilon\|_2^2 + 
\langle \max B^{* \tran} X - \max \hat{B}^\tran X, \epsilon \rangle \right\}
\end{align*}

Putting these two facts back into the original inequality, we have:

\begin{align*}
\frac{1}{2n} \left\{ \| \max B^{* \tran} X - \max \hat{B}^\tran X  \|_2^2 + 
\langle \max B^{* \tran} X - \max \hat{B}^\tran X, \epsilon \rangle \right\} \leq \lambda \| B^* \|_{2,1} - \lambda \|\hat{B}\|_{2,1}
\end{align*}

Let us consider the cross term first. We write $[\max B^{* \tran} X]_i = B^{*\tran}_i X^i$, where $B^*_i$ is a column of the $B^*$ matrix.

Then we have, through two applications of Holder's inequality
\begin{align*}
\frac{1}{2n} |\langle \max B^{* \tran} X - \max \hat{B}^\tran X, \epsilon \rangle| 
				&\leq \frac{1}{2n} \sum_{i=1}^n \epsilon_i \sum_{s=1}^p (B^*_{si} - \hat{B}_{si}) X^i(s) \\
				&\leq \sum_{s=1}^p \frac{1}{2n} \sum_{i=1}^n ( \epsilon_i X^i(s))(B^*_{si} - \hat{B}_{si}) \\
				&\leq \sum_{s=1}^p \| B^*_{s \cdot} - \hat{B}_{s \cdot} \|_2 \frac{1}{2n} \sqrt{\sum_{i=1}^n \epsilon^2_i X^{i 2}(s)} \\
				&\leq \| B^* - \hat{B}\|_{2,1} \max_{s=1,..,p} \frac{1}{2n} \sqrt{\sum_{i=1}^n \epsilon^2_i X^{i 2}(s)} 
\end{align*}

Let us temporarily impose the strong condition that the $\epsilon_i$'s are iid and that $X^i \sim N(0,I_p)$, then the second term $\frac{1}{2n} \sqrt{\sum_{i=1}^n \epsilon^2_i X^{i 2}(s)}  \leq \sigma \sqrt{\frac{\log p}{n}}$ with high probability.

Returning to the main bound, we have:
\begin{align*}
\frac{1}{2n} \left\{ \| \max B^{* \tran} X - \max \hat{B}^\tran X  \|_2^2 + 
\langle \max B^{* \tran} X - \max \hat{B}^\tran X, \epsilon \rangle \right\} &\leq \lambda \| B^* \|_{2,1} - \lambda \|\hat{B}\|_{2,1} 
\end{align*}
Shuffling the terms, we have
\begin{align*}
\frac{1}{2n} \| \max B^{* \tran} X - \max \hat{B}^\tran X  \|_2^2 &\leq \lambda \| B^* \|_{2,1} - \lambda \|\hat{B}\|_{2,1}  + \sigma \sqrt{\frac{\log p}{n}} \| B^* - \hat{B} \|_{2,1} 
\end{align*}

Say $\frac{\lambda}{2} \equiv \sigma \sqrt{\frac{\log p}{n}}$. Now we break apart the norms. Let $S \subset \{1,...,p\}$ be the set of relevant entries of $B^*$. Let $B^*_S$ be a submatrix of $B^*$ where \emph{we restrict the rows} to only $S$--so $B^*_S$ is a

FACT 1: $ \| B^* \|_{2,1} - \|\hat{B}\|_{2,1} \leq \| B^* - \hat{B}_S \|_{2,1} + \| \hat{B}_{S^c} \|_{2,1}$

FACT 2: $\| B^* - \hat{B} \|_{2,1} = \| B^* - \hat{B}_S \|_{2,1} + \| \hat{B}_{S^c} \|_{2,1}$

These two facts combined give us an \textbf{important} intermediate result:
\begin{align}
\frac{1}{2n}  \| \max B^{* \tran} X - \max \hat{B}^\tran X  \|_2^2  &\leq \lambda \| B^* \|_{2,1} - \lambda \|\hat{B}\|_{2,1}  + \frac{\lambda}{2} \| B^* - \hat{B} \|_{2,1}  \notag \\
			&\leq \frac{3\lambda}{2} \| B^* - \hat{B}_S \|_{2,1} - \frac{\lambda}{2} \| \hat{B}_{S^c} \|_{2,1} \label{eqn:intermediate}
\end{align}

In normal Lasso derivation, we would bound the RHS with the restricted eigenvalue (RE) condition. Let us make the following \textbf{assumption}, which is like the RE condition:\\

\begin{framed}
\textbf{The KEY Assumption:} Let $\{ B^*_i \}$ be $s$-sparse vectors.
\begin{align*}
 & \frac{1}{2n} \sum_{i=1}^n ((B_i - B^*_i)^\tran X^i)^2 \geq \kappa^2 \| B_S - B^* \|^2_F \,\,;\:\,\forall B \in \mathcal{B} \\ 
			& \mathcal{B} \equiv \{ B \in \R^{(p+1) \times n} \,:\, (B_i - B_j)^\tran X_i \geq 0 \forall i,j ; \|B_{S^c}\|_{2,1} \leq 3 \| B_S \|_{2,1}\}
\end{align*}
$\kappa$ is some constant, hopefully strictly positive, and hopefully does not decrease too quickly with $n, p$
\end{framed}

Let us optimistically continue to tame the inequality:
\begin{align*}
\frac{1}{2n}  \| \max B^{* \tran} X - \max \hat{B}^\tran X  \|_2^2 &\leq \frac{3\lambda}{2} \| B^* - \hat{B}_S \|_{2,1} - \frac{\lambda}{2} \| \hat{B}_{S^c} \|_{2,1}\\
				&\leq \frac{3 \lambda}{2} \sqrt{|S|} \| B^* - \hat{B}_S \|_F - \frac{\lambda}{2} \|\hat{B}_{S^c} \|_{2,1} \\
				&\leq \frac{3 \lambda}{2}\sqrt{\frac{1}{2n}} \frac{\sqrt{|S|}}{\kappa} \| \max B^{* \tran} X - \max \hat{B}^\tran X \|_2 - \frac{\lambda}{2} \|\hat{B}_{S^c} \|_{2,1}
\end{align*}

Reshuffling and discarding unnecessary terms, we get
\begin{align*}
\sqrt{\frac{1}{2n}} \| \max B^{* \tran} X - \max \hat{B}^\tran X  \|_2  &\leq \frac{3\lambda}{2} \sqrt{|S|}  \frac{1}{\kappa} \\
					&\leq \frac{3}{2} \sqrt{ \frac{ |S| \log p}{n}} \frac{1}{\kappa}
\end{align*}

\textbf{Conclusion 1: } We can denoise well. The estimated convex function $\hat{g}$ and the true convex function $g^*$ agree on the observed $X^i$'s.

Going back to Equation~\ref{eqn:intermediate} and the KEY assumption, we can derive other results:

\[
\| \hat{B}_S - B^* \|_F^2 \leq \frac{3}{2} \sqrt{\frac{|S| \log p}{n} } \frac{1}{\kappa^2}
\]

\textbf{Conclusion 2: } Our estimated convex function $\hat{g}$ agrees with $g^*$ on the relevant dimensions.

\begin{align*}
\| \hat{B}_{S^c} \|_{2,1} & \leq 3 \| \hat{B}_S - B^* \|_{2,1} \\
				&\leq 3 \sqrt{\frac{1}{2n}} \| \max B^{*\tran} X - \max \hat{B}^\tran X \|_2\\
				&\leq \frac{9}{2} \sqrt{\frac{|S|\log p}{n}} \frac{1}{\kappa}
\end{align*}

\textbf{Conclusion 3: } Our estimated regression function is small on the irrelevant dimensions.\\

The next step of the analysis is to bound the interpolation error.

\begin{framed}
The Lipschitz assumption. (can be proven) We suppose that $\hat{g}$ and $g^*$ are both $L$-Lipschitz on a compact subset of $\R^{|S|}$, which is the space of the relevant dimensions; that implies that they must both be $L$-Lipschitz on a compact subset of $\R^{p+1}$ as well.\\

\textbf{Note: } We might only be able to prove a weaker condition described below; but that weaker condition may be sufficient. 
\[
|\hat{g}(x) - \hat{g}(x') | \leq L \| x _S - x'_S \|_2 + O\left( \sqrt{\frac{|S| \log p}{n}} \right)
\]
\end{framed}

Now we can begin to bound the interpolation error. Let $x$ be an arbitrary point in the compact set on which $\hat{g}, g^*$ are Lipschitz and let $X^i$ be the sample in our training data that is closest to $x$.
\begin{align*}
\E|\hat{g}(x) - g^*(x) | & \leq \E| \hat{g}(x) - \hat{g}(X^i) + \hat{g}(X^i) - g^*(X^i) + g^*(X^i) - g^*(x) | \\
		&\leq 2L\E\| x_S - X^i_S \|_2 + O\left( \sqrt{\frac{|S|\log p}{n} } \right)
\end{align*}

For many distributions, $\E\| x_S - X^i_S \|_2$ is on the order of $n^{- \frac{1}{s}}$.

\subsection{Intuition on the KEY Assumption}

Let us suppose $B^* = 0$ for now. Then the KEY assumption becomes:
\begin{align*}
 & \frac{1}{2n} \sum_{i=1}^n (B_i^\tran X^i)^2 \geq \kappa^2 \|B_S \|_F^2 \,\,;\:\,\forall B \in \mathcal{B} \\ 
			& \mathcal{B} \equiv \{ B \in \R^{(p+1) \times n} \,:\, (B_i - B_j)^\tran X_i \geq 0\forall i,j ; \|B_{S^c}\|_{2,1} \leq 3 \| B_S \|_{2,1} ; \}
\end{align*}

\section{Looking at Subgradient}

The linear inequality case:
\begin{align*}
\min_{\beta}  & \|\beta\|_1 \\
\trm{s.t.} & X\beta \leq y
\end{align*}

The Lagrangian optimization is
\[
\max_w \min_\beta \|\beta\|_1 + w^\tran (X\beta - y) 
\]

The KKT condition is that $- w^\tran X \in \partial \|\beta\|_1$ and that $w^\tran X\beta = w^\tran y$.

$w$ is a vector in $\R^n$ and the dual optimization is:
\begin{align*}
\min_w & w^\tran y \\
\trm{s.t.} & X^\tran w \in [-1,1]^p \\
&	w \geq 0
\end{align*}

FACT: Suppose I increase the penalty to $\lambda$--a large number--on the irrelevant coordinates, then the dual becomes
\begin{align*}
\min_w & w^\tran y \\
\trm{s.t.} & X_S^\tran w \in [-1,1]^s \\
& X_{S^c}^\tran w \in [-\lambda, \lambda]^{p-s} \\
&	w \geq 0
\end{align*}

FACT: if an oracle lights up the true support, then the dual becomes:
\begin{align*}
\min_w & w^\tran y \\
\trm{s.t.} & X_S^\tran w \in [-1,1]^s\\
& w \geq 0
\end{align*}
This has strictly fewer constraints than full dimensional problem.

Lastly, we can also compare against\\
\begin{align*}
\min_w & w^\tran X_S \beta^* \\
\trm{s.t.} & X_S^\tran w \in [-1,1]^s\\
& w \geq 0
\end{align*}

Let us consider an easy pair of problems:
\begin{align*}
\min_{\beta}  & \|\beta\|_1 \\
\trm{s.t.} & M \beta \leq M \beta^*
\end{align*}

\begin{align*}
\min_{\beta_S}  & \|\beta_S\|_1 \\
\trm{s.t.} & M_S \beta_S \leq M_S \beta^*_S
\end{align*}
where $M_S$ is a subset of the columns of $M$.

Note that $M \beta = M_S \beta_S + M_{S^c} \beta_{S^c}$.

\section{Using Second-order Information}

\end{document}