### 2013-09-29 tex/scam/nips13/sparsistency.tex

\section{Analysis of Variable Selection Consistency}

We divide our analysis into two parts. We first establish a sufficient
\emph{deterministic} condition for sparsistency.  We then consider the
stochastic setting and argue that the deterministic conditions hold with high probability.

\subsection{Deterministic Setting}

We follow Wainwright \cite{Wain:09a} and define a purely theoretical construct called \emph{restricted regression}.
\begin{definition}
In \emph{restricted regression}, we restrict the indices $k$ in
optimization (\ref{opt:alternate_opt}) to lie in the support $S$ instead of ranging from $1,...,p$. 
\end{definition}

Our analysis then diverges from the now-standard ``primal-dual witness
technique''~\cite{Wain:09a}. Primal-dual witness explicitly solves all the dual variables. Because our optimization is more complex, we do not solve the dual variables on $S$, we instead write the dual variables on $S^c$ as a function of the restricted regression \emph{residual}, which is implicitly a function of the dual variables on $S$.

\begin{theorem} (Deterministic setting)
\label{thm:deterministic}
Let $\{\hat{d}_k, \hat{c}_k \}_{k \in S}$ be the minimizer of the restricted regression, that is, the solution to optimization (\ref{opt:alternate_opt}) where we restrict $k \in S$. Let $\hat{d}_k = 0$ and $ \hat{c}_k = 0$ for $k \in S^c$.
Let $\hat{r} \coloneqq Y - \sum_{k \in S} (\Delta_k \hat{d}_k -
\hat{c}_k \mathbf{1})$ be the restricted regression residual. For $k
\in \{1,...,p\}$, $\Delta_{k, j}$, the $j$-th column of $\Delta_k$, is the $n$-dimensional vector $\max( X_k - X_{k (j)} \mathbf{1}, 0)$. 
Suppose for all $j$ and all $k\in S^c$, $\lambda_n > | \frac{1}{n}
\hat{r}^\tran \Delta_{k,j}|$. Then $\hat{d}_k, \hat{c}_k$ for $k=1,...,p$ is an optimal solution to the full regression \ref{opt:alternate_opt}. Furthermore, any solution to the optimization program \ref{opt:alternate_opt} must be zero on $S^c$.
\end{theorem}
This result holds regardless of whether we impose the boundness and Lipschitz conditions in optimization~\ref{opt:alternate_opt}.
The full proof of Theorem~\ref{thm:deterministic} is in Section~\ref{sec:deterministic_proof} of the Appendix.

\begin{remark}
  The incoherence condition of \cite{Wain:09a} is implicitly encoded
  in our condition on $\lambda_n, \hat{r}, \Delta_{k,j}$. We can
  reconstruct the incoherence condition if we assume that the true
  function $f_0$ is linear and that our fitted functions $\hat{f}_k$
  are linear as well.
\end{remark}

Theorem~\ref{thm:deterministic} allows us to analyze false negative
rates and false positive rates separately. To control false positives,
we study when the condition $\lambda_n > | \frac{1}{n} \hat{r}^\tran
\Delta_{k,j}|$ is fulfilled for all $j$ and all $k \in S^c$. To
control false negatives, we study the restricted regression.

\subsection{Probabilistic Setting}

We use the following statistical setting:

\begin{packed_enum}
\item Let $\mathcal{X} = [-b, b]^p$. Let $F$ be a distribution supported and positive on $\mathcal{X}$. Let $X^{(1)},..., X^{(n)} $ be independent samples from $F$. 
\item Let $Y' = f_0(X) + \epsilon$ where $f_0$ is the true function and $\epsilon$ is noise.  Given $n$ samples $Y'^{(1)},...,Y'^{(n)}$, we input into our optimization $Y = Y' - \bar{Y}'$.
\item Let $S \subset \{1,...,p\}$ denote the relevant variables, i.e.,
  $f_0(X) = f_0(X_S)$, with $s \coloneqq |S|$. 
\end{packed_enum}

Each of our statistical theorems will use a subset of the following assumptions:
\begin{packed_enum}
\item[A1:] $X_S, X_{S^c}$ are independent.  \ A1': $\{ X_k \}_{k \in S}$ are independent.
\item[A2:] $\|f_0\|_\infty \leq B$. \  A2': $f_0$ is convex,
  twice-differentiable, $L$-Lipschitz, and $\trm{supp}(f_0) = S$.
\item[A3:] Suppose $\epsilon$ is mean-zero sub-Gaussian, independent of $X$, with sub-Gaussian scale $\sigma$.
\end{packed_enum}

We will use assumptions A1, A2, A3 to control the probability of false positives and the stronger assumptions A1', A2', A3 to control the probability of false negatives.

\begin{remark}
  We make strong assumptions on the covariates in A1 in order to make
  very weak assumptions on the true regression function $f_0$ in
  A2. In particular, we do not assume that $f_0$ is additive. Relaxing
  these assumptions is an interesting direction for future work.
  %Strong assumptions on the covariates are not uncommon in nonparametric
  %variable selection analysis~\cite{lafferty2008rodeo}. 
  %[TODO: refer to correlated design experiment].
\end{remark}

\begin{theorem} (Controlling false positives) 
\label{thm:false_positive}
Suppose assumptions A1, A2, A3 hold. Suppose also that we run optimization~\eqref{opt:alternate_opt} with the $B$-boundness constraint. Let $c,C$ be absolute constants.
Suppose $\lambda_n \geq c b (sB + \sigma) \sqrt{ \frac{s}{n} \log n
  \log (pn)}$.  Then with probability at least $ 1 - \frac{C}{n}$, for all $j,k$, $\lambda_n >  | \frac{1}{n} \hat{r}^\tran \Delta_{k,j}|$.
Therefore, any solution to the full regression (\ref{opt:alternate_opt}), with boundedness constraint, is zero on $S^c$. 
\end{theorem}

The proof of Theorem~\ref{thm:false_positive} exploits independence of
$\hat{r}$ and $\Delta_{k,j}$ from A1, and then uses concentration of
measure results to argue that $| \frac{1}{n} \hat{r}^\tran
\Delta_{k,j}|$ concentrates around zero at a desired rate. The fact
that $\hat{r}$ is a centered vector is crucial to our proof, and our
theory thus further illustrates the importance of imposing the
centering constraints in optimization \eqref{opt:alternate_opt}. Our
proof uses the concentration of the average of
data sampled \emph{without} replacement
\cite{serfling1974probability}, illustrating that the proof method is not a
trivial application of existing techniques. The full proof of
Theorem~\ref{thm:false_positive} is in
Section~\ref{sec:false_positive_proof} of the Appendix.

\begin{theorem} (Controlling false negatives)
\label{thm:false_negative}
Suppose assumptions A1', A2', A3 hold. Let $\ds \hat{f} = \{ \hat{d}_k, \hat{c}_k\}_{k\in S}$ be any solution to the restricted regression with both the $B$-boundedness and $L$-Lipschitz constraint. Let $c,C$ be absolute constants.
Suppose $L \max \left(\lambda_n, b (B+\sigma)B\sigma \sqrt{\frac{s^5}{n^{4/5}} \log sn} \right) \rightarrow 0$.
Then, for sufficiently large $n$, $\hat{f}_k = (\hat{d}_k, \hat{c}_k)
\neq 0$ for all $k \in S$ with probability at least $1-\frac{C}{n}$.
\end{theorem}

This is a finite sample version of
Theorem~\ref{thm:convex_faithful}. We need stronger assumptions in
Theorem~\ref{thm:false_negative} to use our additive faithfulness
result, Theorem~\ref{thm:convex_faithful}. We also include an extra
Lipschitz constraint so that we can use existing covering number
results \cite{Bronshtein:76}. Recent work
\cite{Guntu:13} shows that the Lipschitz constraint
is not required with more advanced empirical process theory
techniques. We give the full proof of Theorem~\ref{thm:false_negative}
in Section~\ref{sec:false_negative_proof} of the Appendix.

Combining Theorem~\ref{thm:false_positive} and
~\ref{thm:false_negative} and ignoring dependencies on $b,B,L,\sigma$,
we have the following result.
\begin{corollary}
  Assume A1', A2', A3. Let $\lambda_n = \Theta\left( \sqrt{
  \frac{s^3}{n} \log n \log (pn)} \right)$. Suppose $\lambda_n
  \rightarrow 0$ and $\sqrt{\frac{s^5}{n^{4/5}} \log sn} \rightarrow
  0$. Let $\hat{f_n}$ be a solution to (\ref{opt:alternate_opt}) with
  boundedness and Lipschitz constraints. Then 
  $\P( \trm{supp}(\hat{f_n}) = \trm{supp}(f_0) ) \rightarrow 1$.
\end{corollary}
The above corollary implies that sparsistency is achievable at the same exponential scaling of the ambient dimension $p = O(\exp(n^c)), c<1$ rate as parametric models. The cost of nonparametric modeling is reflected in the scaling with respect to $s$, which can only scale at $o(n^{4/25})$.

%\textbf{Comparison with Related Work.}