### 2013-09-30 tex/scam/icml14/paper.tex

\documentclass{article}
\usepackage{amsthm,amsmath,amsfonts,natbib,mathtools,caption,subcaption,sidecap,custom_math}
\usepackage[hscale=0.7,vscale=0.8]{geometry}

\bibliographystyle{plain}

\title{Variable Selection in Convex Function Estimation}

\author{
Min Xu, Minhua Chen, John Lafferty
}

% ENVIRONMENTS
\numberwithin{equation}{section}
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheoremstyle{remark}{\topsep}{\topsep}%
     {\normalfont}% Body font
     {}           % Indent amount (empty = no indent, \parindent = para indent)
     {\bfseries}  % Thm head font
     {.}          % Punctuation after thm head
     {.5em}       % Space after thm head (\newline = linebreak)
     {\thmname{#1}\thmnumber{ #2}\thmnote{ #3}}% Thm head spec
\theoremstyle{remark}
\newtheorem{remark}{Remark}[section]
\newtheorem{example}{Example}[section]
\newtheorem{assumption}{Assumption}[section]
\newtheorem{definition}{Definition}[section]

\begin{document}

% John's macros
\def\X{\mathcal{X}}
\def\comma{\unskip,~}
\def\truep{p^*}
\def\div{\|\,}
\long\def\comment#1{}
\def\reals{{\mathbb R}}
\def\P{{\mathbb P}}
\def\E{{\mathbb E}}
\def\Cov{\mathop{\text{Cov}}}
\def\supp{\mathop{\text{supp}\kern.2ex}}
\def\argmin{\mathop{\text{\rm arg\,min}}}
\def\arginf{\mathop{\text{\rm arg\,inf}}}
\def\argmax{\mathop{\text{\rm arg\,max}}}
\let\tilde\widetilde
\def\csd{${}^*$}
\def\mld{${}^\dag$}
\def\dos{${}^\ddag$}
\def\W{\widetilde Y}
\def\Z{\widetilde X}
\let\hat\widehat
\let\tilde\widetilde
\def\given{{\,|\,}}
\def\ds{\displaystyle}
\def\bs{\backslash}
\def\1{{(1)}}
\def\2{{(2)}}
\def\pn{{(n)}}
\def\ip{{(i)}}
\def\Xbar{\overline{X}}
\def\except{\backslash}
\def\npn{\mathop{\textit{NPN\,}}}
\def\i{{(i)}}
\def\cE{{\mathcal{C}}}
\def\cM{{\mathcal{M}}}
\def\cF{{\mathcal{F}}}
\def\cP{{\mathcal{P}}}
\def\cG{{\mathcal{G}}}
\def\tr{\mathop{\text{tr}}}
\long\def\comment#1{}
\def\ti#1{#1}
\def\titi#1{\textit{#1}}
\def\cram{{\sc cram}}
\def\spam{{\small\sc SpAM}}
\def\diag{\mathop{\rm diag}}
\def\ones{\mathbf{1}}
\def\threebars{\mbox{$|\kern-.25ex|\kern-.25ex|$}}
\def\fatnorm#1{\threebars #1 \threebars}
\def\rank{\mathop{\rm rank}}
\def\S{\mathcal{S}}
\def\H{\mathcal{H}}
\def\K{{K}}
\def\rank{\mathop{\rm rank}}
\def\half{{1/2}}
\def\Y{\mathbb{Y}}
\def\M{\mathbb{M}}
\def\F{\mathbb{F}}
\def\pinv{{-1}}
%\def\ones{\mathds{1}}
%\def\ones{1}
\def\Res{Z}
\def\Proj{P}
\def\cN{{\mathcal N}}
\def\cT{{\mathcal H}}
\def\coloneqq{:=}
\def\mathbf#1{\mbox{\boldmath $#1$}} 
\def\bar#1{\overline{#1}}

\maketitle

\begin{abstract}
  We consider the problem of estimating a sparse convex function of
  many variables.  In constrast to classical nonparametric
  regression with smoothness constraints, we show that convexity is
  additively faithful---it suffices to estimate a convex additive
  model for variable selection.  We develop algorithms for estimating
  sparse convex additive models, including an approach using iterative
  quadratic programming.  Supporting experiments and statistical
  theory are presented, showing variable selection consistency in
  dimensions that can scale exponentially in the sample size.  An
  attractive feature of this framework is the lack of tuning parameters
  for smoothness.
\end{abstract}
\vskip10pt

\input{introduction}
%\input{related_work}
\input{summary}
\input{additive_faithful}
\input{optimization}
\input{sparsistency}
\input{experiments}
\input{discussion}

\clearpage

\bibliography{local}

\newpage
\input{appendix}

\end{document}