### 2013-10-04 tex/scam/icml14/related_work_icml.tex
added section on related work

\section{Related Work}

Variable selection in general nonparametric regression or function
estimation is a notoriously difficult problem.  In even moderate
dimensions, nonparametric regression is subject to the curse
of dimensionality.  But if the function is sparse, depending
on only a few covariates, there is hope that by identifying
the relevant variables one can effectively reduce the 
dimension, and  estimate the function at the
corresponding lower dimensional statistical rate of convergence.

\citet{lafferty2008rodeo} develop a greedy procedure for
adjusting bandwidths in a local linear regression estimator,
and show that the procedure achieves the minimax rate
as if the relevant variables were isolated in advance.
But the method does not carry out explicit variable selection.
Moreover, it only provably scales to dimensions $p$ that 
grow logarithmically in the sample size $n$, so that $p = O(\log n)$.  Note that this
is the opposite of the high dimensional scaling behavior
known to hold for sparsity selection in linear models
using $\ell_1$ penalization, where the sample size $n$
is logarithmic in the dimension $p$. \citet{bertin:08}
develop an optimization-based approach in
the nonparametric setting, applying the lasso
in a local linear model at each test point.  Here again,
however, the method only scales as $p = O(\log n)$,
the low-dimensional regime.
An approximation theory approach to the same
problem is presented in \cite{devore:11}, 
using techniques based on hierarchical hashing schemes,
similar to those used for ``junta'' problems \cite{mossel:04}.
Here it is shown that the sample complexity scales as $n > \log p$ 
if one adaptively selects the points on
which the high-dimensional function is evaluated.
\citet{dalalyan:12} give
support for the intrinsic difficulty of variable
selection in nonparametric regression, giving lower bounds 
showing that sparsistency is not possible if $n < \log p$ or if $n <
\exp s$, where $s$ is the number of relevant variables.
Variable selection over kernel classes is studied
by \citet{Kolch:10}.

Perhaps most closely related to the present work, is the framework
studied by \cite{Raskutti:12} for sparse additive models, where sparse
regression is considered under an additive assumption, with each
component function belonging to an RKHS.  An advantage of working over
an RKHS, in contrast to the other papers mentioned above, is that
regression with a sparsity-inducing regularization penalty can be
formulated as a finite dimensional convex cone optimization.
On the other hand, 
smoothing parameters for the component Hilbert spaces
must be chosen, leading to extra tuning parameters
that are difficult to select in practice.  In addition, 
the additive model must be assumed to
be correct for sparsistent variable selection.

An attraction of the convex function estimation
framework we consider in this paper
is that the additive model can be
used for convenience, without assuming it
to actually hold.  Moreover, our estimator has
no tuning parameters.  As we show below,
our method scales to high dimensions, with
a dependence on the intrinsic dimension $s$ 
that scales polynomially, rather than exponentially
as in the general case analyzed in \cite{dalalyan:12}.

%* Laetitia Comminges, Arnak S. Dalalyan 
%Tight conditions for consistency of variable selection 
%In the context of high dimensionality. 
%Ann. Statist, 40(5), 2667-2696, 2012 
%
%Comment: They estimate the coefficient of the regression function with respect to a Fourier basis and then threshold the estimated coefficients. They prove that sparsistency is achievable so long as $n > log p$ and $n > exp(s)$. They assume that the functions are smooth with respect to the Fourier basis. The fourier coefficients however seem difficult to estimate; it appears that the knowledge of the true density is required. They also provide 
%
%
%* Bertin, K. and Lecu´e, G. (2008). Selection of variables 
%and dimension reduction in high-dimensional non-parametric 
%regression. Electron. J. Stat. 2 1224–1241 
%
%Comment: They assume Holder smoothness and use multiple L1 regularized local polynomial smoothing to detect the sparsity pattern. They require that $p < log n$.
%
%
%* "Approximation of Functions of Few Variables in High Dimensions" 
%by DeVore, Petrova and Wojtaszczyk. 
%
%Comment: They show that variable selection under a scaling of
%