### 2014-04-18 tex/scam/estimation.tex
paper update

\section{Estimation Procedure}
\label{sec:acdc}

Theorem~\ref{thm:acdc_faithful} suggest a two stage procedure for variable selection in convex regression--first learning a sparse convex additive model and then, on the residual, separately learning an univariate concave function for each of the dimensions.

More precisely, given samples $(\mathbf{x}_1, y_1), ..., (\mathbf{x}_n, y_n)$, we perform the following procedure, which we refer to as AC/DC (additively convex / decoupled concave):
\begin{enumerate}
\item Compute: 
\begin{equation}
\label{eqn:scam}
\hat{f}_1, ..., \hat{f}_p = \argmin_{f_1,...,f_p \in \mathcal{C}^1} 
   \frac{1}{n} \sum_{i=1}^n \Big(y_i - \sum_{k=1}^p f_k(x_{ki}) \Big)^2 
       + \lambda \sum_{k=1}^p \| f_k \|_\infty
\end{equation}
\item Compute, for each $k$ such that $\| \hat{f}_k \| = 0$:
\begin{equation}
\label{eqn:dc}
\hat{g}_k = \argmin_{g_k \in \mathcal{C}^1} 
   \frac{1}{n} \sum_{i=1}^n \Big( y_i - \sum_{k'} \hat{f}_{k'}(x_{k'i}) 
    - g_k(x_{ki})\Big)^2 
      + \lambda \| g_k \|_\infty
\end{equation}
\item Output as the set of relevant variables: 
$\{ k \,:\, \| \hat{f}_k \|_\infty > 0 
  \textrm{ or } \| \hat{g}_k \|_\infty > 0 \}$ 
\end{enumerate}

We added an $\ell_\infty/\ell_1$ penalty in equation~\ref{eqn:scam} and the $\ell_\infty$ penalty in equation~\ref{eqn:dc} to encourage sparsisty. There are other forms of penalty that can also produce sparse estimates such as a penalty on the derivative of each of the component functions; we find the form we use to be convenient both for theoretical analysis and implementation.

% DO NOT CHANGE; RefTex variables -minx

%%% Local Variables: ***
%%% mode:latex ***
%%% TeX-master: "paper.tex" ***
%%% End: ***