### 2014-06-29 tex/scam/abstract.tex
added abstract.tex

\begin{abstract}
  We study the problem of variable selection in convex nonparametric
  regression.  Under the assumption that the true regression function
  is convex and sparse, we develop a screening procedure to select a
  subset of variables that contains the relevant variables. Our
  approach is a two-stage quadratic programming method that estimates
  a sum of one-dimensional convex functions, followed by
  one-dimensional concave regression fits on the residuals.  In
  contrast to previous methods for sparse additive models, the
  optimization is finite dimensional and requires no tuning parameters
  for smoothness. Under appropriate assumptions, we prove that the
  procedure is faithful in the population setting, yielding no false
  negatives, and we give a finite sample statistical analysis. In
  addition, we introduce algorithms for efficiently carrying out the
  required quadratic programs. The approach leads to significant
  computational and statistical advantages over fitting a full model,
  and provides an effective, practical approach to variable screening
  in convex regression.
\end{abstract}

\begin{keyword}
\kwd{nonparametric regression}
\kwd{convex regression}
\kwd{variable selection}
\kwd{quadratic programming}
\kwd{additive model}
\end{keyword}