### 2014-07-01 tex/scam/introduction.tex
restructured related work

\section{Introduction}

Shape restrictions such as monotonicity, convexity, and concavity
provide a natural way of limiting the complexity of many statistical
estimation problem.  Shape-constrained estimation is not as well
understood as more traditional nonparametric estimation involving
smoothness constraints.  For instance, as of this writing, the minimax
rate of convergence for multivariate convex regression has yet to be
rigorously established in full generality.  Even the one-dimensional
case is challenging, and has been of recent interest
\citep{guntusen:13}.

In this paper we study the problem of variable selection in convex
regression.  Assuming that the regression function is convex and
sparse, our goal is to identify the relevant variables.  We show that
it suffices to estimate a sum of $p$ one-dimensional convex functions,
leading to significant computational and statistical advantages.  This
is in contrast to general nonparametric regression, where fitting an
additive model can result in false negatives.  Our approach is based
on a two-stage quadratic programming procedure.  In the first stage,
we fit an convex additive model, imposing a sparsity penalty.  In the
second stage, we fit a concave function on the residual for each
variable.  As we demonstrate, this perhaps non-intuity second stage is
in general necessary.  Our first result is that this procedure is
faithful in the population setting, meaning that it results in no
false negatives, under mild assumptions on the density of the
covariates.  Our second result is a finite-sample statistical analysis
of the procedure, where we upper bound the statistical rate of
convergence.  Each stage of the method requires quadratic programming.
An additional contribution is to show how these quadratic programs can
be forumlated to be more scalable.  Simulations demonstrate the method
performs in a manner that is consistent with our analysis.

Estimation of convex functions arises naturally in several
applications.  Examples include geometric programming \citep{Boyd04},
computed tomography \citep{Prince:90}, target reconstruction
\citep{Lele:92}, image analysis \citep{Golden:06} and circuit design
\citep{Hannah:12}.  Other applications include queuing theory
\citep{Chen:01} and economics, where it is of interest to estimate
concave utility functions \citep{Pratt:68}.  See \cite{Lim:12} for
other applications.  
Beyond cases where the assumption of convexity is
natural, the convexity assumption can be attractive as a
tractable, nonparamametric relaxation of the linear model.

Variable selection in general nonparametric regression or function
estimation is a notoriously difficult problem. \citet{lafferty2008rodeo} develop a greedy procedure for
adjusting bandwidths in a local linear regression estimator,
and show that the procedure achieves the minimax rate
as if the relevant variables were isolated in advance.
But the method only provably scales to dimensions $p$ that 
grow logarithmically in the sample size $n$, so that $p = O(\log n)$.  Note that this
is the opposite of the high dimensional scaling behavior
known to hold for sparsity selection in linear models
using $\ell_1$ penalization, where the sample size $n$
is logarithmic in the dimension $p$. \citet{bertin:08}
develop an optimization-based approach in
the nonparametric setting, applying the lasso
in a local linear model at each test point.  Here again,
however, the method only scales as $p = O(\log n)$,
the low-dimensional regime.
An approximation theory approach to the same
problem is presented in \cite{devore:11}, 
using techniques based on hierarchical hashing schemes,
similar to those used for ``junta'' problems \cite{mossel:04}.
Here it is shown that the sample complexity scales as $n > \log p$ 
if one adaptively selects the points on
which the high-dimensional function is evaluated.

\citet{dalalyan:12} show that the exponential scaling $p=O(\log n)$ is achievable if the underlying function is assumed to be smooth with respect to a Fourier basis. They also give
support for the intrinsic difficulty of variable
selection in nonparametric regression, giving lower bounds 
showing that sparsistency is not possible if $n < \log p$ or if $n <
\exp s$, where $s$ is the number of relevant variables.
Variable selection over kernel classes is studied
by \citet{Kolch:10}.

Perhaps most closely related to the present work, is the framework
studied by \cite{Raskutti:12} for sparse additive models, where sparse
regression is considered under an additive assumption, with each
component function belonging to an RKHS.  An advantage of working over
an RKHS, in contrast to the other papers mentioned above, is that
nonparametric regression with a sparsity-inducing regularization penalty can be
formulated as a finite dimensional convex cone optimization.
On the other hand, 
smoothing parameters for the component Hilbert spaces
must be chosen, leading to extra tuning parameters
that are difficult to select in practice.  In addition, 
the additive model must be assumed to
be correct for sparsistent variable selection.

An attraction of the convex function estimation framework we consider
in this paper is that the additive model can be used for convenience,
without assuming it to actually hold.  While nonparametric, the
problem is naturally formulated using finite dimensional convex
optimization, but with no additional tuning parameters.  As we show
below, our method scales to high dimensions, with a dependence on the
intrinsic dimension $s$ that scales polynomially, rather than
exponentially as in the general case analyzed in \cite{dalalyan:12}.

%* Laetitia Comminges, Arnak S. Dalalyan 
%Tight conditions for consistency of variable selection 
%In the context of high dimensionality. 
%Ann. Statist, 40(5), 2667-2696, 2012 
%
%Comment: They estimate the coefficient of the regression function with respect to a Fourier basis and then threshold the estimated coefficients. They prove that sparsistency is achievable so long as $n > log p$ and $n > exp(s)$. They assume that the functions are smooth with respect to the Fourier basis. The fourier coefficients however seem difficult to estimate; it appears that the knowledge of the true density is required. They also provide 
%
%
%* Bertin, K. and Lecu´e, G. (2008). Selection of variables 
%and dimension reduction in high-dimensional non-parametric 
%regression. Electron. J. Stat. 2 1224–1241 
%
%Comment: They assume Holder smoothness and use multiple L1 regularized local polynomial smoothing to detect the sparsity pattern. They require that $p < log n$.
%
%
%* "Approximation of Functions of Few Variables in High Dimensions" 
%by DeVore, Petrova and Wojtaszczyk. 
%
%Comment: They show that variable selection under a scaling of
%

%Variable selection for nonparametric
%regression under smoothness constraints is difficult without
%making additional strong assumptions. 
%\cite{lafferty2008rodeo} show that computationally
%efficient, near minimax-optimal estimation is possible, but in ambient
%dimensions that scale only as $p = O(\log n)$; see also \cite{BertinLecue}.  This is in contrast
%to the $p=O\bigl(e^{n^c}\bigr)$ scaling enjoyed by sparse linear
%models \cite{Wain:09a}. \citet{dalalyan:12} achieve exponential scaling $p=O(e^n)$
%under certain Fourier smoothness conditions, but require
%that the number of relevant variables $s$ must be less than $\log n$.
%
%Approximating the regression function by a sum of one-dimensional
%functions, known as sparse additive models \citep{Ravikumar:09}, is a
%practical alternative to fully nonparametric function estimation.  
%Sparse additive models have been further studied in by \cite{Meier09},
%and by \cite{Raskutti:12}.  But 
%the additive assumption is limited.  In particular, the natural idea
%of first selecting the single variable effects, then the pairwise
%effects, and so on, does not in general lead to consistent variable
%selection.  In other words, the general nonparametric model is not
%additively faithful.  Remarkably, the additional assumption of
%convexity does lead to consistent variable selection, as we show
%here. In addition, we show that the high dimensional scaling $n =
%O\big(\textrm{poly}(s) \log p\big)$ is achievable for sparse convex
%additive models. Thus, with respect to variable selection, the
%geometric convexity constraint is quite different from the smoothness
%constraints imposed in traditional nonparametric regression.
%

In the following section we give a high-level summary of our technical
results, including additive faithfulness, variable selection 
consistency, and high dimensional scaling.  In Section~X...

% DO NOT CHANGE; RefTex variables -minx

%%% Local Variables: ***
%%% mode:latex ***
%%% TeX-master: "paper.tex" ***
%%% End: ***