### 2014-07-07 tex/scam/summary.tex
major edits

\def\x{\mathbf{x}}

\section{Overview of Results}

In this section we provide a high-level description of our technical
results.  Our presentation is intended to give the reader
an overview of our results and their significance.  The 
full technical details, the precise statement of the results,
and their detailed proofs are provided in following sections.

Our results address three problems.  The first is to develop
a procedure based on an additive approximation for identifying
relevant variables in convex regression.  To this end,
we prove a result that shows when and how the additive approximation
be used without introducing false negatives in the population setting.  The
second result concerns the efficient implementation of
the quadratic programs required by the procedure.  
Finally, the third problem we consider is the finite-sample
behavior of our proposed method.

We first establish some notation, to be used throughout the remainder 
of the paper.
If $\mathbf{x}$ is a vector, we use $\mathbf{x}_{-k}$ to denote the
vector with the $k$-th coordinate removed. If $\mathbf{v} \in \R^n$, then
$v_{(1)}$ denotes the smallest coordinate of $\mathbf{v}$ in
magnitude, and $v_{(j)}$ denotes the $j$-th smallest; $\mathbf{1}_n \in \R^n$
is the all ones vector. If $X \in \R^p$ and $S \subset
\{1,...,p\}$, then $X_S$ is the subvector of $X$ restricted to
the coordinates in $S$. Given $n$ samples $X^{(1)},...,X^{(n)}$, we use
$\bar{X}$ to denote the empirical average. Given a random variable
$X_k$ and a scalar $x_k$, we use $\E[\cdot \given x_k]$ as a shorthand
for $\E[ \cdot \given X_k = x_k]$.

\subsection{Additive faithfulness}

The starting point for our approach is the observation that least squares
nonparametric estimation under convexity constraints is equivalent to
a finite dimensional quadratic program.  Specifically, the infinite
dimensional optimization 
\begin{align}
\begin{split}
\text{minimize} & \quad \sum_{i=1}^n (Y_i - f(\x_i))^2 \\
\text{subject to} &  \quad f:\reals^p\rightarrow\reals\ \text{is
  convex}
\end{split}
\end{align}
is equivalent to the finite dimensional quadratic
program 
\begin{align}
\begin{split}
\text{minimize}_{f, \beta} & \;\; \sum_{i=1}^n (Y_i - f_i)^2 \\
\text{subject to} & \;\; f_j \geq f_i + \beta_i^T (\x_j-\x_i),\; \text{for
    all $i,j$}.
\end{split}
\end{align}
See \cite{Boyd04}, Section 6.5.5.
Here $f_i$ is the estimated function value $f(\x_i)$, and the vectors
$\beta_i \in \reals^d$ represent supporting hyperplanes to the
epigraph of $f$.  Importantly, this finite dimensional quadratic program does
not have tuning parameters for smoothing the function.

For general regression, using an additive approximation for variable
selection may make errors.  In particular, the nonlinearities in the
regression function may result in an additive component being wrongly
zeroed out.  Surprisingly, this will not happen for convex regression
under appropriate conditions.

A differential function $f$ depends on variable $x_k$ if
$\partial_{x_k} f \neq 0$ with probability greater than zero.  An additive approximation is given by
\begin{equation}
\{f_k^*\}, \mu^* \coloneqq \argmin_{f_1,\ldots, f_p, \mu} \Bigl\{ 
             \E ( f(X) - \mu - \sum_{k=1}^p f_k(X_k))^2 
         \,:\, \E f_k(X_k) = 0 \Bigr\}.
\end{equation}
We say that $f$ is \textit{additively faithful} in case $f^*_k = 0$
implies that $f$ does not depend on coordinate $k$.
% We can define the support $\trm{supp}(f) \coloneqq \{ k \,:\,
% \trm{$k$ is relevant to $f$}\}$. Let $f^* = \sum_{k=1}^s$, then $f$
% is additively faith if $\trm{supp}(f) = \trm{supp}(f^*)$.
Additive faithfulness is a desireable property,
since it implies that an additive approximation may allow us to 
screen out irrelevant variables.

Our first result shows that convex multivariate functions are
additively faithful
under the following assumption on the distribution of the data.
\begin{definition}
  Let $p(\mathbf{x})$ be a density supported on $[0,1]^p$.  Then $p$
  satisfies the \emph{boundary-points condition} if, for all $j$, and
  for all $\mathbf{x}_{-j}$:
\[
\frac{\partial p(\mathbf{x}_{-j} \given x_j)}{\partial x_j}  =  
\frac{\partial^2 p(\mathbf{x}_{-j} \given x_j)}{\partial x_j^2} = 0
\quad \trm{at $x_j = 0$ and $x_j = 1$}.
\]
\end{definition}

As discussed in Section~\ref{sec:additivefaithful}, this is a relatively weak condition. 
Our first result is that this condition suffices in the population
setting of convex regression.

\begin{theorem}
  Let $p$ be a positive density supported on $C=[0,1]^s$ that
  satisfies the boundary-points property. If $f$ is convex and twice
  differentiable, then $f$ is additively faithful under $p$.
\end{theorem}

% We give the full proof in Section~\ref{sec:faithful_proof} of the
% Appendix, but pause here to provide some intuition.

Intuitively, an additive approximation zeroes out variable $k$ when, fixing $x_k$, every
``slice'' of $f$ integrates to zero. We prove this result
by showing that ``slices'' of convex
functions that integrate to zero cannot be ``glued together'' while
still maintaining convexity.

While this shows that convex functions are additively faithful, it is difficult to
estimate the optimal additive functions $f^*_k$'s.  The difficulty
is that $f^*_k$ need not be
a convex function, as we show through a counterexample
in Section~\ref{sec:additivefaithful}.   
Since the true regression function $f$ is convex, it is natural to ask
when it is sufficient to estimate an convex additive model.  
Unfortunately, a convex additive approximation is not generally
faithful.  In other words, it could be that $f^*_k\equiv 0$
even for a relevant variable $x_k$.  But our next result
shows that this type of error can be detected by fitting 
a \textit{concave} function to the residual.  If the
concave fit to the residual is zero, we can then safely
mark $x_k$ as an irrelevant variable.

\begin{theorem}
Suppose $p(\mathbf{x})$ is a positive
density on $C=[0,1]^p$ and satisfies the boundary-points
condition. Suppose that $f$ is convex and twice-differentiable.
and that $\partial_{x_k} f$, $\partial_{x_k} p( \mathbf{x}_{-k}
\given x_k )$, and $\partial_{x_k}^2 p( \mathbf{x}_{-k} \given x_k)$
are all continuous as functions on $C$.\\
Define
\begin{equation}
\{ f^*_k \}_{k=1}^p,\mu^* = \arg\min \Big \{
\E\big( f(X) - \mu - \sum_{k=1}^s f_k(X_k) \big)^2 \,:\, f_k \in
\mathcal{C}^1, \, \E f_k(X_k) = 0 \Big \}
\end{equation} where $\mathcal{C}^1$ is the set of univariate convex
functions, and
\begin{equation}
g^*_k = \arg\min \Big\{ \E\big( f(X) -
\sum_{k' \neq k} f^*_{k'}(X_{k'}) - g_k \big)^2 \,:\, g_k \in \mh
\mathcal{C}^1, \E g_k(X_k) = 0 \Big\},
\end{equation}
with $\mh{}\mathcal{C}^1$ denoting the set of univariate concave
function.  Then, $f^*_k = 0$ and $g^*_k = 0$ implies that $f$ does not
depend on $x_k$, i.e., $\partial_{x_k} f(\mathbf{x}) = 0$ with
probability one.
\end{theorem}

This result suggests a two-stage screening
procedure for variable selection. In the first stage we fit a sparse \emph{convex}
additive model.  In the second stage we
fit a concave function $\hat g_k$ to the residual for each variable
having a zero convex component $\hat f_k$.  If both $\hat f_k = 0$ and
$\hat g_k = 0$, we remove variable $x_k$.  
Our next results concern the optimizations involved, and their finite
sample statistical performance.

\subsection{Optimization}

\subsection{Finite-sample analysis}