### 2014-07-11 tex/scam/summary.tex
added section on finite sample analysis

\def\x{\mathbf{x}}

\section{Overview of Results}

In this section we provide a high-level description of our technical
results.  Our presentation is intended to give the reader
an overview of our results and their significance.  The 
full technical details, the precise statement of the results,
and their detailed proofs are provided in following sections.

Our results address three problems.  The first is to develop
a procedure based on an additive approximation for identifying
relevant variables in convex regression.  To this end,
we prove a result that shows when and how the additive approximation
be used without introducing false negatives in the population setting.  The
second result concerns the efficient implementation of
the quadratic programs required by the procedure.  
Finally, the third problem we consider is the finite sample
behavior of our proposed method.

We first establish some notation, to be used throughout the remainder 
of the paper.
If $\mathbf{x}$ is a vector, we use $\mathbf{x}_{-k}$ to denote the
vector with the $k$-th coordinate removed. If $\mathbf{v} \in \R^n$, then
$v_{(1)}$ denotes the smallest coordinate of $\mathbf{v}$ in
magnitude, and $v_{(j)}$ denotes the $j$-th smallest; $\mathbf{1}_n \in \R^n$
is the all ones vector. If $X \in \R^p$ and $S \subset
\{1,...,p\}$, then $X_S$ is the subvector of $X$ restricted to
the coordinates in $S$. Given $n$ samples $X^{(1)},...,X^{(n)}$, we use
$\bar{X}$ to denote the empirical average. Given a random variable
$X_k$ and a scalar $x_k$, we use $\E[\cdot \given x_k]$ as a shorthand
for $\E[ \cdot \given X_k = x_k]$.

\subsection{Faithful screening}

The starting point for our approach is the observation that least squares
nonparametric estimation under convexity constraints is equivalent to
a finite dimensional quadratic program.  Specifically, the infinite
dimensional optimization 
\begin{align}
\begin{split}
\text{minimize} & \quad \sum_{i=1}^n (Y_i - f(\x_i))^2 \\
\text{subject to} &  \quad f:\reals^p\rightarrow\reals\ \text{is
  convex}
\end{split}
\end{align}
is equivalent to the finite dimensional quadratic
program 
\begin{align}
\label{eq:convreg}
\begin{split}
\text{minimize}_{f, \beta} & \;\; \sum_{i=1}^n (Y_i - f_i)^2 \\
\text{subject to} & \;\; f_j \geq f_i + \beta_i^T (\x_j-\x_i),\; \text{for
    all $i,j$}.
\end{split}
\end{align}
See \cite{Boyd04}, Section 6.5.5.
Here $f_i$ is the estimated function value $f(\x_i)$, and the vectors
$\beta_i \in \reals^d$ represent supporting hyperplanes to the
epigraph of $f$.  Importantly, this finite dimensional quadratic program does
not have tuning parameters for smoothing the function.

For general regression, using an additive approximation for variable
selection may make errors.  In particular, the nonlinearities in the
regression function may result in an additive component being wrongly
zeroed out.  Surprisingly, this will not happen for convex regression
under appropriate conditions.

A differential function $f$ depends on variable $x_k$ if
$\partial_{x_k} f \neq 0$ with probability greater than zero.  An additive approximation is given by
\begin{equation}
\{f_k^*\}, \mu^* \coloneqq \argmin_{f_1,\ldots, f_p, \mu} \Bigl\{ 
             \E ( f(X) - \mu - \sum_{k=1}^p f_k(X_k))^2 
         \,:\, \E f_k(X_k) = 0 \Bigr\}.
\end{equation}
We say that $f$ is \textit{additively faithful} in case $f^*_k = 0$
implies that $f$ does not depend on coordinate $k$.
% We can define the support $\trm{supp}(f) \coloneqq \{ k \,:\,
% \trm{$k$ is relevant to $f$}\}$. Let $f^* = \sum_{k=1}^s$, then $f$
% is additively faith if $\trm{supp}(f) = \trm{supp}(f^*)$.
Additive faithfulness is a desireable property,
since it implies that an additive approximation may allow us to 
screen out irrelevant variables.

Our first result shows that convex multivariate functions are
additively faithful
under the following assumption on the distribution of the data.
\begin{definition}
  Let $p(\mathbf{x})$ be a density supported on $[0,1]^p$.  Then $p$
  satisfies the \emph{boundary-points condition} if, for all $j$, and
  for all $\mathbf{x}_{-j}$:
\[
\frac{\partial p(\mathbf{x}_{-j} \given x_j)}{\partial x_j}  =  
\frac{\partial^2 p(\mathbf{x}_{-j} \given x_j)}{\partial x_j^2} = 0
\quad \trm{at $x_j = 0$ and $x_j = 1$}.
\]
\end{definition}

As discussed in Section~\ref{sec:additivefaithful}, this is a relatively weak condition. 
Our first result is that this condition suffices in the population
setting of convex regression.

\begin{theorem}
  Let $p$ be a positive density supported on $C=[0,1]^s$ that
  satisfies the boundary-points property. If $f$ is convex and twice
  differentiable, then $f$ is additively faithful under $p$.
\end{theorem}

% We give the full proof in Section~\ref{sec:faithful_proof} of the
% Appendix, but pause here to provide some intuition.

Intuitively, an additive approximation zeroes out variable $k$ when, fixing $x_k$, every
``slice'' of $f$ integrates to zero. We prove this result
by showing that ``slices'' of convex
functions that integrate to zero cannot be ``glued together'' while
still maintaining convexity.

While this shows that convex functions are additively faithful, it is difficult to
estimate the optimal additive functions $f^*_k$'s.  The difficulty
is that $f^*_k$ need not be
a convex function, as we show through a counterexample
in Section~\ref{sec:additivefaithful}.   
Since the true regression function $f$ is convex, it is natural to ask
when it is sufficient to estimate an convex additive model.  
Unfortunately, a convex additive approximation is not generally
faithful.  In other words, it could be that $f^*_k\equiv 0$
even for a relevant variable $x_k$.  But our next result
shows that this type of error can be detected by fitting 
a \textit{concave} function to the residual.  If the
concave fit to the residual is zero, we can then safely
mark $x_k$ as an irrelevant variable.

\begin{theorem}
Suppose $p(\mathbf{x})$ is a positive
density on $C=[0,1]^p$ and satisfies the boundary-points
condition. Suppose that $f$ is convex and twice-differentiable.
and that $\partial_{x_k} f$, $\partial_{x_k} p( \mathbf{x}_{-k}
\given x_k )$, and $\partial_{x_k}^2 p( \mathbf{x}_{-k} \given x_k)$
are all continuous as functions on $C$.
Define
\begin{equation}
\{ f^*_k \}_{k=1}^p,\mu^* = \arg\min \Big \{
\E\big( f(X) - \mu - \sum_{k=1}^s f_k(X_k) \big)^2 \,:\, f_k \in
\mathcal{C}^1, \, \E f_k(X_k) = 0 \Big \}
\end{equation} where $\mathcal{C}^1$ is the set of univariate convex
functions, and
\begin{equation}
g^*_k = \arg\min \Big\{ \E\big( f(X) -
\sum_{k' \neq k} f^*_{k'}(X_{k'}) - g_k \big)^2 \,:\, g_k \in \mh
\mathcal{C}^1, \E g_k(X_k) = 0 \Big\},
\end{equation}
with $\mh{}\mathcal{C}^1$ denoting the set of univariate concave
functions.  Then $f^*_k = 0$ and $g^*_k = 0$ implies that $f$ does not
depend on $x_k$, i.e., $\partial_{x_k} f(\mathbf{x}) = 0$ with
probability one.
\end{theorem}

This result naturally suggests a two-stage screening
procedure for variable selection. In the first stage we fit a sparse convex
additive model.  In the second stage we
fit a concave function $\hat g_k$ to the residual for each variable
having a zero convex component $\hat f_k$.  If both $\hat f_k = 0$ and
$\hat g_k = 0$, we can safely discard variable $x_k$.  
Our next results concern the required optimizations, and their finite
sample statistical performance.

\subsection{Optimization}

In Section~\ref{sec:optimization} we present optimization algorithms for the additive convex regression stage.
The convex constraints for the additive functions, analogous to 
the multivariate constraints \eqref{eq:convreg},
are  that each component $f_{k}(\cdot)$ 
can be represented by its supporting hyperplanes, i.e.,
\begin{equation}
      f_{ki'} \geq f_{ki} + \beta_{ki}(x_{ki'}-x_{ki}) \quad \text{for
        all $i,i'$}
\end{equation}
where $f_{ki}\coloneqq f_{k}(x_{ki})$ and $\beta_{ki}$ is the
subgradient at point $x_{ki}$. While this apparently requires $O(n^2
p)$ equations to impose the supporting hyperplane constraints, 
in fact, only $O(np)$ constraints suffice.  This is because univariate convex functions are
characterized by the condition that the subgradient, which is a scalar, must
increase monotonically. This observation leads to a reduced quadratic
program with $O(np)$ variables and $O(np)$ constraints.

Directly applying a QP solver to this optimization is still computationally
expensive for relatively large
$n$ and $p$.  We thus develop a block
coordinate descent method, where in each step we solve a sparse
QP involving $O(n)$ variables and $O(n)$ constraints.  This 
is efficiently solved using optimization packages 
such as {\sc mosek}.  The details of these optimizations
are given in Section~\ref{sec:optimization}.

\subsection{Finite sample analysis}

In Section~\ref{sec:finitesample} 
we analyze the finite sample variable selection consistency of convex
additive modeling, without making the assumption that the true
regression function is additive.  Our analysis first establishes
a sufficient deterministic condition for variable selection 
consistency, and then considers a stochastic setting.
Our proof technique decomposes the KKT conditions for the optimization
in a manner that is similar to the now standard \emph{primal-dual
  witness} method~\cite{wainwright2009sharp}.

We prove separate results that allow us to analyze false negative
rates and false positive rates.  To control false positives,
we analyze when the condition on $\lambda_n$ is fulfilled for all $j$
and all $k \in S^c$, where $S\subset \{1,\ldots, p\}$ is the set of
relevant variables. 
To control false negatives, we analyze the restricted regression
where the variables in $S^c$ are zeroed out, following the primal-dual
strategy.

Each of our theorems uses a subset of the following assumptions:
\begin{packed_enum}
\item[A1:] $X_S, X_{S^c}$ are independent. 
\item[A2:] $f_0$ is convex and twice-differentiable. 
\item[A3:] $|\partial_{x_k} f_0 | \leq L$ for all $k$
\item[A4:] The noise $w$ is mean-zero sub-Gaussian, independent of $X$
\end{packed_enum}

We define $\alpha_f, \alpha_g$ as a measure of the signal strength of the weakest variable:
\begin{align*}
\alpha_f &= \inf_{f \in \mathcal{C}^p \,:\, \exists k ,\, f^*_k \neq 0 \,\wedge\, f_k = 0} 
       \Big\{ \mathbb{E} \big( f_0(X) - f(X) \big)^2 - 
        \mathbb{E} \big( f_0(X) - f^*(X) \big)^2  \Big\}\\
\alpha_g &=   \min_{k \in S \,:\, g^*_k \neq 0}
      \Big\{ \mathbb{E} \big( f_0(X) - f^*(X) \big)^2 - 
    \mathbb{E} \big( f_0(X) - f^*(X) - g^*_k(X_k) \big)^2 \Big\}
\end{align*}

Intuitively, if $\alpha_f$ is smaller, then it is easier to make a
false omission in the additive convex stage of the procedure. If
$\alpha_g$ is smaller, then it is easier to make a false omission in
the decoupled concave stage of the procedure.

We make strong assumptions on the covariates in A1 in order to make
very weak assumptions on the true regression function $f_0$ in
A2; in particular, we do not assume that $f_0$ is additive.

We also include an extra
Lipschitz constraint so that we can use existing covering number
results \cite{Bronshtein:76}. Recent work
\cite{Guntu:13} shows that the Lipschitz constraint
is not required with more advanced empirical process theory
techniques; we leave the incorporation of this development as future work.

Our main result is the following.
\begin{theorem}
Suppose assumptions A1-A4 hold. Let $\hat{f}$ be any AC solution and
$\hat{g}_k$'s be any DC solution, both with $L$-Lipschitz constraints,
both with  regularization parameter $\lambda$ scaling as
$\lambda = \Theta \Big( sLb \sqrt{\frac{1}{n} \log^2 np} \Big)$. 
Suppose in addition that
\begin{gather}
\frac{\alpha_f}{\sigma} \geq c (Lb)^3 \sqrt{\frac{s^5}{n^{4/5}} \log^2
  np}\\
\frac{\alpha_g^2}{\sigma} \geq c (Lb)^3 \sqrt{\frac{s^5}{n^{4/5}}
  \log^2 2np}.
\end{gather}
Then, for sufficiently large $n$, with probability at least $1-\frac{1}{n}$:
\begin{align*}
\hat{f}_k \neq 0 \trm{ or } \hat{g}_k \neq 0 &\trm{ for all } k \in S\\
\hat{f}_k = 0 \trm{ and } \hat{g}_k = 0 & \trm{ for all } k \notin S.
\end{align*}

\end{theorem}

This shows that variable selection consistency is achievable at the
same exponential scaling of the ambient dimension $p = O(\exp(n^c)),
c<1$ rate,  as for linear models. The cost of nonparametric modeling is
reflected in the scaling with respect to $s$, which can only scale at
$o(n^{4/25})$.

We remark that \citet{dalalyan:12} show that under tradtional smoothness
constraints, variable selection is achievable only if $n > O(e^s)$. 
Here we demonstrate that convexity yields the scaling $n =
O(\textrm{poly}(s))$.