### 2014-07-11 tex/scam/faithful.tex

\documentclass[12pt,pdftex,aos]{imsart}
%\documentclass[12pt,pdftex,generic,noinfoline]{imsart}

\RequirePackage[OT1]{fontenc}
\usepackage{amsthm,amsmath,amsfonts,natbib,mathtools,custom_math}
\RequirePackage{hypernat}
\usepackage[ruled,section]{algorithm}
\usepackage[noend]{algorithmic}
\usepackage{graphicx}
\usepackage[hscale=0.7,vscale=0.8]{geometry}
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{yhmath}

\renewcommand{\baselinestretch}{1.1}
%\parskip12pt
%\parindent0pt
\setcounter{tocdepth}{2}

\input{macros}

\begin{document}

\begin{frontmatter}
\title{Faithful Variable Screening for High-Dimensional Convex Regression}
\runtitle{Faithful Variable Screening for Convex Regression}

\begin{aug}
\vskip10pt
\author{\fnms{Min} \snm{Xu}\ead[label=e1]{minx@cs.cmu.edu}}
\comma 
\author{\fnms{Minhua} \snm{Chen}\ead[label=e2]{minhua@galton.uchicago.edu}}
\and
\author{\fnms{John} \snm{Lafferty}\ead[label=e3]{lafferty@galton.uchicago.edu}}
\address{
\begin{tabular}{ccc}
\\[10pt]
Machine Learning Department & Department of Statistics & Department of Statistics\\
Carnegie Mellon University  & University of Chicago & Department of Computer Science\\
& & University of Chicago
\end{tabular}
%\\[20pt]
%\printead{e1,e2,e3}
}
\end{aug}

\input{abstract}
%\tableofcontents
\vskip20pt
\end{frontmatter}

\maketitle
\vskip10pt

%\tableofcontents

\input{introduction}
\input{summary}
\input{additive_faithful}
\input{estimation}
\input{optimization}
\input{sparsistency}
\input{experiments}
\input{discussion}

\clearpage

\newpage
\input{appendix}
%\newpage
%\input{notes}

\section*{Acknowledgements}
Research supported in part by NSF grants IIS-1116730, 
AFOSR grant FA9550-09-1-0373, ONR grant
N000141210762, and an Amazon AWS in Education Machine Learning
Research grant.

\bibliographystyle{imsart-nameyear}
\bibliography{local}

\end{document}