### 2014-11-09 tex/scam/coverletter.tex
edit

\documentclass[12pt]{article}
\input{letterhead}
\recommendee{}

\begin{document}
\thispagestyle{empty}
\myletterhead{}
\date{}
\vskip15pt

To the Editors:

We would like to submit the manuscript ``Faithful variable screening
for high-dimensional convex regression'' for review and possible
publication in {\it The Annals of Statistics}.  This paper develops a
new procedure based on additive models for variable selection in high
dimensional convex regression, shows that the procedure yields no
false negatives in the population setting, and gives a finite sample
statistical analysis.

The manuscript is 29 pages including references, and we have
prepared a supplement with full proofs of all of the technical
results.  No version of this paper has been previously
published in any form.

As possible expert reviewers, we suggest Adityanand Guntuboyina
(Berkeley), Richard Samworth (Cambridge), Bodhisattva Sen (Columbia)
and Ming Yuan (Wisconsin).

\jdlucsig{}

\end{document}