### 2015-12-07 tex/scam/paper-submit.tex

\documentclass[12pt,pdftex,aos,noinfoline,addressasfootnote]{imsart}
%\documentclass[12pt,pdftex,generic,noinfoline]{imsart}

\RequirePackage[OT1]{fontenc}
\usepackage{amsthm,amsmath,amsfonts,natbib,mathtools,custom_math,amssymb}
\RequirePackage{hypernat}
\usepackage[ruled,section]{algorithm}
\usepackage[noend]{algorithmic}
\usepackage{graphicx}
%\usepackage[hscale=0.7,vscale=0.8]{geometry}
\usepackage[text={6.0in,8.6in},centering]{geometry}
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{yhmath}
\usepackage[usenames,dvipsnames]{xcolor}
\definecolor{shadecolor}{gray}{0.9}
\definecolor{shadecolor}{gray}{0.9}
\hypersetup{citecolor=MidnightBlue}
\hypersetup{linkcolor=MidnightBlue}
\hypersetup{urlcolor=MidnightBlue}
\usepackage{enumerate}
\usepackage{multirow}

\renewcommand{\baselinestretch}{1.045}
%\parskip12pt
%\parindent0pt
\setcounter{tocdepth}{2}

\input{macros}
\arxiv{1411.1805}

\begin{document}

\begin{frontmatter}
\title{Faithful Variable Screening for\\ High-Dimensional Convex Regression}
\runtitle{Faithful Variable Screening for Convex Regression}

\begin{aug}
\vskip10pt
\author{\fnms{Min} \snm{Xu${}^*$}\ead[label=e1]{minx@cs.cmu.edu}}
\comma 
\author{\fnms{Minhua} \snm{Chen${}^\dag$}\ead[label=e2]{minhua@galton.uchicago.edu}}
\and
\author{\fnms{John}
  \snm{Lafferty${}^\dag$}\ead[label=e3]{lafferty@galton.uchicago.edu}}
\address{${}^*$Machine Learning Department, Carnegie Mellon University,
  Pittsburgh PA 15213}
\address{${}^\dag$Department of Statistics, University of Chicago, Chicago IL 60637}
\end{aug}

\input{abstract} 
%\tableofcontents
\vskip20pt 
\end{frontmatter}

\maketitle

\vskip10pt

%\tableofcontents

\input{introduction}
\input{summary}
\input{additive_faithful}
\input{estimation}
\input{optimization}
\input{sparsistency}
\input{experiments}
\input{discussion}
%\newpage
%\input{notes}

\section*{Acknowledgements}
Research supported in part by NSF grants IIS-1116730 and DMS-1513594,
AFOSR grant FA9550-09-1-0373, ONR grant N000141210762, and an Amazon
AWS in Education Machine Learning Research grant. The authors 
thank  Ming Yuan for valuable discussions and suggestions
on this research. The authors are also grateful to the anonymous
reviewers whose careful proofreading and helpful comments significantly
improved the quality of the paper.

%\clearpage
\bibliographystyle{imsart-nameyear}
\bibliography{local}

%\clearpage
\input{supplement}

\end{document}