### 2016-03-08 tex/scam/faithful.tex
edits for final version

\documentclass[pdftex,aos]{imsart}

\RequirePackage[OT1]{fontenc}
\RequirePackage{amsthm,amsmath}
\usepackage{amsfonts,mathtools,custom_math,amssymb}
\RequirePackage[numbers]{natbib}
\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{multirow}
\usepackage{xr}
\externaldocument{supplement}

% settings
%\pubyear{2005}
%\volume{0}
%\issue{0}
%\firstpage{1}
%\lastpage{8}
\arxiv{arXiv:0000.0000}

\startlocaldefs
\numberwithin{equation}{section}
\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\endlocaldefs

\input{macros}

\begin{document}

\begin{frontmatter}
\title{Faithful Variable Screening for High-Dimensional Convex Regression}
\runtitle{Faithful Variable Screening for Convex Regression}

\begin{aug}
\author{\fnms{Min} \snm{Xu}\thanksref{t1,m1}\ead[label=e1]{min.cut@gmail.com}},
\author{\fnms{Minhua} \snm{Chen}\thanksref{m2}\ead[label=e2]{minhua@somewhere.com}}
\and
\author{\fnms{John} \snm{Lafferty}\thanksref{t1,m3}\ead[label=e3]{lafferty@galton.uchicago.edu}}

\thankstext{t1}{Research supported in part by NSF grants IIS-1116730 and DMS-1513594,
AFOSR grant FA9550-09-1-0373, ONR grant N00014-12-10762, and an Amazon
AWS in Education Machine Learning Research grant.}
\runauthor{M.~Xu, M.~Chen and J.~Lafferty}

\affiliation{University of Pennsylvania\thanksmark{m1}, Amazon.com\thanksmark{m2}, and University of Chicago\thanksmark{m3}}

\address{Address of M.~Xu\\
Department of Statistics\\
The Wharton School\\
University of Pennsylvania\\
Philadelphia, Pennsylvania 19104\\
USA\\
\printead{e1}\\
}

\address{Address of M.~Chen\\
Amazon.com\\
410 Terry Avenue North\\
Seattle, Washington 98109\\
USA\\
\printead{e2}\\}

\address{Address of J.~Lafferty\\
Department of Statistics\\
University of Chicago\\
5734 South University Avenue\\
Chicago, Illinois 60637\\
USA\\
\printead{e3}\\}

\end{aug}

\begin{abstract}
  We study the problem of variable selection in convex nonparametric
  regression.  Under the assumption that the true regression function
  is convex and sparse, we develop a screening procedure to select a
  subset of variables that contains the relevant variables. Our
  approach is a two-stage quadratic programming method that estimates
  a sum of one-dimensional convex functions, followed by
  one-dimensional concave regression fits on the residuals.  In
  contrast to previous methods for sparse additive models, the
  optimization is finite dimensional and requires no tuning parameters
  for smoothness. Under appropriate assumptions, we prove that the
  procedure is faithful in the population setting, yielding no false
  negatives.  We give a finite sample statistical analysis, and
  introduce algorithms for efficiently carrying out the required
  quadratic programs. The approach leads to computational and
  statistical advantages over fitting a full model, and provides an
  effective, practical approach to variable screening in convex
  regression.
\end{abstract}

\begin{keyword}[class=MSC]
\kwd[Primary ]{62G08}
\kwd[; secondary ]{52A41}
\end{keyword}

\begin{keyword}
\kwd{nonparametric regression}
\kwd{convex regression}
\kwd{variable selection}
\kwd{quadratic programming}
\kwd{additive model}
\end{keyword}

\end{frontmatter}

\input{introduction}
\input{summary}
\input{additive_faithful}
\input{estimation}
\input{optimization}
\input{sparsistency}
\input{experiments}
\input{discussion}

\section*{Acknowledgements}
The authors thank  Ming Yuan for valuable discussions and suggestions
on this research. The authors are also grateful to the anonymous
reviewers whose careful proofreading and helpful comments significantly
improved the quality of the paper.

\begin{supplement}[id=suppA]
%\sname{Supplement A}
\stitle{Supplement to Faithful Variable Screening for High-Dimensional Convex Regression}
\slink[doi]{doi}
\sdatatype{.pdf}
\sdescription{The supplement provides detailed proofs of certain technical results,
together with further explanation of the Gaussian example
and simplifications when the density is a product.}
\end{supplement}

%\clearpage
\bibliographystyle{imsart-nameyear}
\bibliography{local}

\end{document}